---
title: "Wright Mediation Framework: Effect size metric, Variance Decomposition, and Power/Sensitivity Analysis "
subtitle: "A Unified Approach for Complex Mediation Designs"
author: "mediationgourp"
date: "2026-06-24"
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    theme: cosmo
    highlight-style: github
  typst:
    toc: true
    toc-depth: 3
    toc-title: "Index"
    number-sections: true
---

## 1. Introduction to the Effect Size Metric

### 1.1 The Need for a Standardized Effect Size Metric in Mediation

In structural equation modeling (SEM) and mediation analysis, quantifying the magnitude of an indirect effect relies primarily on the product of standardized path coefficients ($\beta_a \cdot \beta_b$). Although this approach standardizes the effect by expressing it in standard deviation units, the resulting product functions as a metric of relative linear change rather than a bounded proportion of shared variance. Consequently, it does not directly reflect the operational contribution of the mediation pathway to the total variance of the outcome variable. This limitation makes it difficult to evaluate the practical significance of the mechanism within an absolute, variance-accounted-for framework that may be directly comparable to traditional effect sizes.

### 1.2 Sewall Wright’s Path Tracing Rules

To derive a formalized effect size, this framework leverages Sewall Wright’s (1934) seminal path tracing rules. These foundational principles state that the total variance of an outcome variable can be exactly decomposed into the sum of the structural pathways connecting the variable to itself, either through direct, indirect, or common causal tracks.

In a standard single-mediator structure where an independent variable ($X$) affects the dependent variable ($Y$) directly via path $c'$ and indirectly through a mediator ($M$) via paths $a$ and $b$, Wright's rules partition the total observed variance of the outcome ($\sigma^2_Y$) into the following additive components:

$$\sigma^2_Y = {c'}^2 \sigma^2_X + a^2 b^2 \sigma^2_X + 2 a b c' \sigma^2_X + b^2 \sigma^2_{e_M} + \sigma^2_{e_Y}$$

Where: \* ${c'}^2 \sigma^2_X$ represents the variance uniquely explained by the direct effect of $X$. \* $a^2 b^2 \sigma^2_X$ represents the variance uniquely explained by the indirect pathway ($X \rightarrow M \rightarrow Y$). \* $2 a b c' \sigma^2_X$ captures the unanalyzed covariance between the direct and indirect paths arising from their shared upstream cause ($X$). \* $b^2 \sigma^2_{e_M}$ represents the variance transmitted to $Y$ from the unexplained residual of the mediator ($e_M$). \* $\sigma^2_{e_Y}$ is the residual error variance intrinsic to $Y$.

By systematically accounting for every linear component, this exact decomposition provides the mathematical foundation necessary to isolate the unique variance explained ($R^2$) by the indirect pathway without confounding or double-counting effects across multi-stage structures.

### 1.3 The Coefficient of Determination ($R^2$) as an Effect Size

Building upon Wright's decomposition, this framework proposes evaluating the indirect effect as a strict proportion of the outcome's variance. By expressing the structural mechanism as a Coefficient of Determination ($R^2_{\text{indirect}}$), the indirect effect is mapped onto a bounded scalar ranging from $0$ to $1$. This provides a standardized metric representing the specific amount of variance the mediation pathway explains within the final outcome. By anchoring the indirect effect to a bounded $0$ to $1$ scale, this variance-based approach resolves the interpretative ambiguities inherent in unbounded standardized products. The subsequent section formalizes how this conceptual framework is translated into a precise, computable mathematical model.

------------------------------------------------------------------------

## 2. Theoretical Foundations of $R^2_{\text{indirect}}$

### 2.1 Mathematical Formulation

The `r2_indirect` function calculates the proportion of variance explained by specific indirect paths. For a single mediation model defined by unstandardized path coefficients $a$ and $b$, the variance of the independent variable $\sigma^2_X$, and the variance of the dependent variable $\sigma^2_Y$, the metric is algebraically formulated as:

$$R^2_{\text{indirect}} = \frac{(a \cdot b)^2 \cdot \sigma^2_X}{\sigma^2_Y}$$

### 2.2 Algorithmic Implementation

The custom R function extracts the unstandardized point estimates and the implied covariance matrix directly from a fitted structural model. It computes the squared product of the targeted indirect paths and scales it by the ratio of the predictor's variance to the outcome's variance. To provide a measure of precision and statistical inference, the function utilizes a non-parametric bootstrapping procedure (`n_boot`) to compute 95% Confidence Intervals for the estimated variance proportion. Ultimately, this algorithmic procedure ensures that the theoretical variance proportion is not only accurately extracted from the covariance matrix but also bounded by robust inferential metrics. To guarantee the structural integrity of these point estimates, it is necessary to mathematically prove the underlying variance decomposition, a validation addressed in the following section.

------------------------------------------------------------------------

## 3. Algebraic Validation of the Metric (`wright_proof`)

To validate the proposed metric algebraically, the framework includes the `wright_proof` suite. This validation ensures that the variance components extracted from the structural paths map accurately onto the theoretically implied variance of the outcome variable, guaranteeing mathematical consistency.

### 3.1 Mathematical Derivation of the 5-Block Decomposition

Consider a simple mediation model defined by the following structural equations, where all variables are expressed as deviations from their means:

$$M = aX + e_M$$ $$Y = c'X + bM + e_Y$$

Under standard ordinary least squares (OLS) and SEM assumptions, the independent variable $X$, the mediator residual $e_M$, and the outcome residual $e_Y$ are mutually independent:

($\text{Cov}(X, e_M) = \text{Cov}(X, e_Y) = \text{Cov}(e_M, e_Y) = 0$)

Substituting the equation for $M$ into the equation for $Y$ isolates the independent sources of variance:

$$Y = c'X + b(aX + e_M) + e_Y$$ $$Y = (c' + ab)X + be_M + e_Y$$

Applying the variance operator ($\text{Var}$) to both sides of the equation yields:

$$\text{Var}(Y) = \text{Var}\left((c' + ab)X\right) + \text{Var}(be_M) + \text{Var}(e_Y)$$

Extracting the constant coefficients from the variance operator results in:

$$\sigma^2_Y = (c' + ab)^2\sigma^2_X + b^2\sigma^2_{eM} + \sigma^2_{eY}$$

Expanding the binomial term $(c' + ab)^2 = (c')^2 + (ab)^2 + 2c'ab$ provides the exact 5-block decomposition implemented in the verification script:

$$\sigma^2_Y = (c')^2\sigma^2_X + (ab)^2\sigma^2_X + 2c'ab\sigma^2_X + b^2\sigma^2_{eM} + \sigma^2_{eY}$$

Where the five distinct components are defined as: 1. **Pure Direct Effect:** $(c')^2 \cdot \sigma^2_X$ 2. **Pure Indirect Effect:** $(a \cdot b)^2 \cdot \sigma^2_X$ 3. **Joint Covariance:** $2 \cdot c' \cdot a \cdot b \cdot \sigma^2_X$ 4. **Independent Mediator Error:** $b^2 \cdot \sigma^2_{eM}$ 5. **Residual Outcome Error:** $\sigma^2_{eY}$

### 3.2 Statistical Justification of the Joint Covariance Term

The inclusion of the third component, $2c'ab\sigma^2_X$, is required by the probability theorem for the variance of the sum of two random variables. For any two random variables $A$ and $B$:

$$\text{Var}(A + B) = \text{Var}(A) + \text{Var}(B) + 2\cdot\text{Cov}(A,B)$$

In this structural context, let $A = c'X$ and $B = abX$. The covariance between these two linear combinations is derived as:

$$\text{Cov}(c'X, abX) = c' \cdot ab \cdot \text{Cov}(X, X) = c'ab\sigma^2_X$$

Therefore, the joint covariance term is an essential algebraic component of the total variance of $Y$. Omitting this term would violate principles of linear algebra and result in an incorrect reconstruction of the observed variance.

### 3.3 Interpretation Under Suppression Conditions

The decomposition accounts for statistical suppression effects through the sign of the Joint Covariance block. When the direct effect ($c'$) and the indirect effect ($a \cdot b$) have opposing signs, the product $c'ab$ is negative, causing the entire joint covariance term to become negative ($2c'ab\sigma^2_X < 0$).

In this scenario, the negative term subtracts from the total implied variance, modeling how the opposing pathways offset one another. To maintain the algebraic equality with the total observed variance $\sigma^2_Y$, the residual outcome error variance ($\sigma^2_{eY}$) must expand. This mathematically reflects the operational reality of suppression designs, where opposing paths reduce the overall explained variance of the model.

### 3.4 Scaling the Decomposition for Complex Models

The algebraic validation scales to accommodate more complex mediation architectures. For instance:

- **Two Parallel Mediators:** The variance of the outcome variable is partitioned into 10 distinct blocks. This structure accounts for the unanalyzed residual covariance between parallel mediators ($2 \cdot b_1 \cdot b_2 \cdot \text{Cov}(e_{M1}, e_{M2})$) and the multi-way joint structural covariances.
- **Two Serial Mediators:** The structure expands to 15 distinct components, tracking the sequential variance flow through the internal path connecting the mediators ($d_{21}$).

In all configurations, the `wright_proof` function verifies that the final residual error block remains non-negative and that the total decomposed variance precisely matches the observed variance of the outcome. This procedure may be extended to even more complex models. This scalable algebraic proof confirms that the proposed variance decomposition maintains strict mathematical consistency regardless of the model's structural configuration or the presence of suppression dynamics. Having established this theoretical foundation for variance decomposition, the framework now transitions to evaluating the formal boundaries of the indirect effect size index. The following section develops the mathematical conditions governing the boundedness of $R^2_{\text{indirect}}$, seamlessly bridging analytical proofs with systematic empirical validation

------------------------------------------------------------------------

## 4. Boundedness of $R^2_{\text{indirect}}$: Formal Theorem and Empirical Validation

```{r}
#| label: setup
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)

theme_set(
  theme_minimal(base_size = 12) +
    theme(
      plot.title    = element_text(face = "bold", size = 13),
      plot.subtitle = element_text(size = 11, color = "grey40"),
      axis.title    = element_text(size = 11)
    )
)
```

### 4.1 Index Definition

As previously specified, for a simple mediation model with unstandardized path coefficients $a$, $b$, $c'$, predictor variance $\sigma^2_X$, and outcome variance $\sigma^2_Y$, the index is defined as:

$$R^2_{\text{indirect}} = \frac{(a \cdot b)^2 \cdot \sigma^2_X}{\sigma^2_Y}$$

The total variance of the outcome decomposes exactly into five components (Wright, 1934):

$$\sigma^2_Y = \underbrace{(c')^2\sigma^2_X}_{\text{pure direct}} +
              \underbrace{(ab)^2\sigma^2_X}_{\text{pure indirect}} +
              \underbrace{2c'ab\sigma^2_X}_{\text{joint covariance}} +
              \underbrace{b^2\sigma^2_{e_M}}_{\text{mediator error}} +
              \underbrace{\sigma^2_{e_Y}}_{\text{outcome error}}$$

### 4.2 Theorem: Boundedness Condition

### Theorem 1 (Necessary and Sufficient Condition for $R^2_{\text{indirect}} \leq 1$)

Let a simple mediation model be given with structural equations:

$$M = aX + e_M, \qquad Y = c'X + bM + e_Y$$

under standard OLS assumptions ($\text{Cov}(X, e_M) = \text{Cov}(X, e_Y) =
\text{Cov}(e_M, e_Y) = 0$). Then:

$$R^2_{\text{indirect}} \leq 1 \iff (c')^2 + 2c'ab \geq
-\frac{b^2\sigma^2_{e_M} + \sigma^2_{e_Y}}{\sigma^2_X}$$

In an equivalent form, the condition is satisfied **if and only if**:

$$(c' + ab)^2\sigma^2_X + b^2\sigma^2_{e_M} + \sigma^2_{e_Y} \geq (ab)^2\sigma^2_X$$

that is, $\sigma^2_Y \geq (ab)^2\sigma^2_X$.

### 4.3 Formal Proof

**Step 1 — Expression of the condition.**

By definition, $R^2_{\text{indirect}} \leq 1$ is equivalent to:

$$\frac{(ab)^2\sigma^2_X}{\sigma^2_Y} \leq 1
\iff (ab)^2\sigma^2_X \leq \sigma^2_Y$$

**Step 2 — Substitution of the decomposition of** $\sigma^2_Y$.

From the five-block decomposition:

$$\sigma^2_Y = (c' + ab)^2\sigma^2_X + b^2\sigma^2_{e_M} + \sigma^2_{e_Y}$$

The condition becomes:

$$(ab)^2\sigma^2_X \leq (c' + ab)^2\sigma^2_X + b^2\sigma^2_{e_M} + \sigma^2_{e_Y}$$

**Step 3 — Algebraic simplification.**

Subtracting $(ab)^2\sigma^2_X$ from both sides:

$$0 \leq \left[(c' + ab)^2 - (ab)^2\right]\sigma^2_X + b^2\sigma^2_{e_M} + \sigma^2_{e_Y}$$

Expanding the square of the binomial:

$$(c' + ab)^2 - (ab)^2 = (c')^2 + 2c'ab$$

We therefore obtain:

$${0 \leq \left[(c')^2 + 2c'ab\right]\sigma^2_X + b^2\sigma^2_{e_M} + \sigma^2_{e_Y}}$$

Which is equivalent to the condition stated in Theorem 1.

### 4.4 Corollary: Standardized Coefficients

### Corollary 1 (Sufficient Condition in Standardized Terms)

If the standardized path coefficients satisfy $|b^*| \leq 1$, then $R^2_{\text{indirect}} \leq 1$.

**Proof.**

The standardized coefficient of $b$ is defined as:

$$b^* = b \cdot \frac{\sigma_M}{\sigma_Y}$$

From which:

$$b^{*2} = b^2 \cdot \frac{\sigma^2_M}{\sigma^2_Y}
         = b^2 \cdot \frac{a^2\sigma^2_X + \sigma^2_{e_M}}{\sigma^2_Y}$$

The condition $|b^*| \leq 1$ implies $b^2\sigma^2_M \leq \sigma^2_Y$, which means $b^2(a^2\sigma^2_X + \sigma^2_{e_M}) \leq \sigma^2_Y$.

Since $(ab)^2\sigma^2_X \leq b^2(a^2\sigma^2_X + \sigma^2_{e_M})$ (as the second term also includes $b^2\sigma^2_{e_M} \geq 0$), we have:

$$(ab)^2\sigma^2_X \leq b^2\sigma^2_M \leq \sigma^2_Y$$

which is precisely the condition of Theorem 1.

### 4.5 Empirical Validation

### 4.5.1 Grid Search over the Parameter Space

The following code performs a systematic search over a dense grid of parameter values, verifying in what proportion of cases $R^2_{\text{indirect}}$ exceeds 1 and which condition on the standardized coefficients is associated with the violations.

```{r}
#| label: grid-search
#| cache: false

# -----------------------------------------------------------------------
# Main function
# -----------------------------------------------------------------------

compute_r2_indirect <- function(a, b, cp,
                                 var_X  = 1,
                                 var_eM = 1,
                                 var_eY = 1) {
  var_M   <- a^2 * var_X + var_eM
  var_Y   <- (cp + a * b)^2 * var_X + b^2 * var_eM + var_eY
  r2      <- (a * b)^2 * var_X / var_Y
  a_std   <- a * sqrt(var_X) / sqrt(var_M)
  b_std   <- b * sqrt(var_M) / sqrt(var_Y)

  data.frame(
    a = a, b = b, cp = cp,
    var_eM = var_eM, var_eY = var_eY,
    var_Y  = var_Y,
    r2     = r2,
    a_std  = a_std,
    b_std  = b_std,
    exceeds     = r2 > 1,
    suppression = sign(cp) != sign(a * b)
  )
}

# -----------------------------------------------------------------------
# Parameter grid
# -----------------------------------------------------------------------

params <- expand.grid(
  a      = seq(0.1, 3.0, by = 0.2),
  b      = seq(0.1, 3.0, by = 0.2),
  cp     = seq(-3.0, 3.0, by = 0.3),
  var_eM = c(0.1, 0.5, 1.0),
  var_eY = c(0.1, 0.5, 1.0)
)

results <- mapply(
  FUN      = compute_r2_indirect,
  a        = params$a,
  b        = params$b,
  cp       = params$cp,
  var_eM   = params$var_eM,
  var_eY   = params$var_eY,
  SIMPLIFY = FALSE
) |> dplyr::bind_rows()

# -----------------------------------------------------------------------
# Summary
# -----------------------------------------------------------------------

n_total  <- nrow(results)
n_exceed <- sum(results$exceeds)
max_r2   <- max(results$r2)

cat("=================================================================\n")
cat("  BOUNDEDNESS TEST — R²_indirect\n")
cat("=================================================================\n\n")
cat(sprintf("Tested parameter combinations      : %s\n",
            format(n_total, big.mark = ",")))
cat(sprintf("Violations (R²_indirect > 1)       : %s  (%.2f%%)\n",
            format(n_exceed, big.mark = ","),
            100 * n_exceed / n_total))
cat(sprintf("Maximum observed R²_indirect       : %.4f\n\n", max_r2))

supp <- results |> dplyr::filter(suppression)
cat("--- Suppression conditions (cp and ab of opposite signs) ---\n")
cat(sprintf("Total suppression cases            : %s\n",
            format(nrow(supp), big.mark = ",")))
cat(sprintf("Violations under suppression       : %s  (%.2f%%)\n",
            format(sum(supp$exceeds), big.mark = ","),
            100 * mean(supp$exceeds)))
cat(sprintf("Max R²_indirect under suppression  : %.4f\n\n",
            max(supp$r2)))

# Verification of the necessary condition
viol <- results |> dplyr::filter(exceeds)
no_viol_with_small_std <- viol |>
  dplyr::filter(abs(a_std) <= 1, abs(b_std) <= 1)

cat("--- Corollary 1 Verification ---\n")
cat(sprintf("|b*| > 1 in ALL violation cases            : %s\n",
            ifelse(all(abs(viol$b_std) > 1), "CONFIRMED", "FALSIFIED")))
cat(sprintf("Violations with |a*| ≤ 1 AND |b*| ≤ 1      : %d\n",
            nrow(no_viol_with_small_std)))

# Realistic range
realistic <- results |>
  dplyr::filter(abs(a_std) <= 0.6, abs(b_std) <= 0.6)
cat(sprintf("\nRealistic range (|a*|,|b*| ≤ 0.6)\n"))
cat(sprintf("  Cases tested   : %s\n", format(nrow(realistic), big.mark = ",")))
cat(sprintf("  Violations     : %d  (%.2f%%)\n",
            sum(realistic$exceeds), 100 * mean(realistic$exceeds)))
```

### 4.5.2 Plot 1 — Parameter Space Map

The following plot shows $R^2_{\text{indirect}}$ as a function of $a$ and $c'$ (with $b = 1.5$, $\sigma^2_{e_M} = \sigma^2_{e_Y} = 0.5$). The red line indicates the $R^2 = 1$ boundary.

```{r}
#| label: fig-heatmap
#| fig-cap: "Map of R²_indirect in the parameter space (a, c'). The red line indicates the R² = 1 boundary. The bottom-right region corresponds to boundedness violations."
#| fig-width: 8
#| fig-height: 6

# Fine grid for the plot
b_fix  <- 1.5
eM_fix <- 0.5
eY_fix <- 0.5

plot_grid <- expand.grid(
  a  = seq(0.01, 3.0, length.out = 200),
  cp = seq(-3.0, 3.0, length.out = 200)
) |>
  dplyr::mutate(
    b     = b_fix,
    var_Y = (cp + a * b)^2 + b^2 * eM_fix + eY_fix,
    r2    = (a * b)^2 / var_Y,
    r2_vis = pmin(r2, 3)
  )


ggplot(plot_grid, aes(x = a, y = cp)) +
  geom_tile(aes(fill = r2_vis)) +
  geom_contour(aes(z = r2), breaks = 1,
               color = "red", linewidth = 1.2) +
  scale_fill_gradientn(
    colors = c("navy", "steelblue", "white", "orange", "firebrick"),
    values = scales::rescale(c(0, 0.5, 1, 1.5, 3)),
    name   = expression(R^2[indirect] ~ "(cap 3)")
  ) +
  annotate("text", x = 2.5, y = -2.4,
           label = "Violation\nregion",
           color = "red", fontface = "bold", size = 3.8,
           hjust = 0.5) +
  annotate("text", x = 0.8, y = 2.0,
           label = "Always\nbounded",
           color = "white", fontface = "bold", size = 3.8,
           hjust = 0.5) +
  labs(
    title    = expression(R^2[indirect] ~ "in the parameter space"),
    subtitle = bquote(b == .(b_fix) ~~ sigma^2[eM] == .(eM_fix) ~~ sigma^2[eY] == .(eY_fix)),
    x        = "Path coefficient a",
    y        = "Direct effect c'"
  ) -> p_heatmap

print(p_heatmap)
```

### 4.5.3 Plot 2 — Distribution under Suppression

```{r}
#| label: fig-distribution
#| fig-cap: "Distribution of R²_indirect under suppression conditions. The red dashed line indicates the threshold value R² = 1."
#| fig-width: 7
#| fig-height: 5

pct_exceed_supp <- 100 * mean(supp$r2 > 1)

supp |>
  dplyr::mutate(r2_vis = pmin(r2, 6)) |>
  ggplot(aes(x = r2_vis)) +
  geom_histogram(bins = 70, fill = "steelblue",
                 color = "white", alpha = 0.85) +
  geom_vline(xintercept = 1, color = "red",
             linewidth = 1.2, linetype = "dashed") +
  annotate("text", x = 1.15, y = Inf,
           label = expression(R^2 == 1),
           color = "red", hjust = 0, vjust = 1.8, size = 4) +
  labs(
    title    = expression("Distribution of" ~ R^2[indirect] ~ "under suppression conditions"),
    subtitle = sprintf("%.1f%% of cases exceed R² = 1", pct_exceed_supp),
    x        = expression(R^2[indirect] ~ "(truncated at 6)"),
    y        = "Count"
  ) -> p_dist

print(p_dist)
```

### 4.5.4 Plot 3 — Standardized Coefficients and Violations

```{r}
#| label: fig-standardized
#| fig-cap: "Relationship between |b*| and R²_indirect (left) and distribution of the maximum standardized coefficient by violation status (right)."
#| fig-width: 13
#| fig-height: 5

# Sample for the scatter plot
set.seed(42)
sample_df <- results |>
  dplyr::sample_n(min(8000, nrow(results))) |>
  dplyr::mutate(
    stato  = ifelse(exceeds, "Violation (R² > 1)", "Bounded (R² ≤ 1)"),
    r2_vis = pmin(r2, 5),
    max_std = pmax(abs(a_std), abs(b_std))
  )

p1 <- ggplot(sample_df, aes(x = abs(b_std), y = r2_vis, color = stato)) +
  geom_point(alpha = 0.25, size = 1.2) +
  geom_hline(yintercept = 1, color = "red",
             linewidth = 1, linetype = "dashed") +
  geom_vline(xintercept = 1, color = "darkorange",
             linewidth = 1, linetype = "dashed") +
  scale_color_manual(values = c("steelblue", "firebrick")) +
  annotate("text", x = 0.3, y = 4.5,
           label = "|b*| = 1", color = "darkorange",
           angle = 90, size = 3.5, vjust = -0.3) +
  labs(
    title  = expression(R^2[indirect] ~ "vs" ~ "|b*|"),
    x      = expression("|b*| (standardized coefficient b)"),
    y      = expression(R^2[indirect] ~ "(truncated at 5)"),
    color  = NULL
  ) +
  theme(legend.position = "bottom") +
  xlim(0, 7)

results_hist <- results |>
  dplyr::mutate(
    stato   = ifelse(exceeds, "Violation (R² > 1)", "Bounded (R² ≤ 1)"),
    max_std = pmin(pmax(abs(a_std), abs(b_std)), 6)
  )

p2 <- ggplot(results_hist, aes(x = max_std, fill = stato, color = stato)) +
  geom_density(alpha = 0.45, linewidth = 0.8) +
  geom_vline(xintercept = 1, color = "black",
             linewidth = 1.2, linetype = "dashed") +
  scale_fill_manual(values  = c("steelblue", "firebrick")) +
  scale_color_manual(values = c("steelblue", "firebrick")) +
  annotate("rect", xmin = 0, xmax = 0.6,
           ymin = -Inf, ymax = Inf,
           alpha = 0.08, fill = "forestgreen") +
  annotate("text", x = 0.3, y = Inf,
           label = "Realistic\nrange",
           color = "forestgreen", vjust = 1.5, size = 3.2,
           fontface = "italic") +
  labs(
    title  = expression("Distribution of" ~ max("|a*|", "|b*|")),
    x      = expression(max("|a*|", "|b*|") ~ "(truncated at 6)"),
    y      = "Density",
    fill   = NULL, color = NULL
  ) +
  theme(legend.position = "bottom") +
  xlim(0, 6)

print(p1 + p2)
```

### 4.5.5 Plot 4 — Boundary Surface

The following plot shows the region of the $(a, b)$ space where a violation is **mathematically possible** (for some value of $c'$ with an opposite sign to $ab$), with $\sigma^2_{e_M} = \sigma^2_{e_Y} = 0.5$.

```{r}
#| label: fig-boundary
#| fig-cap: "Region of the (a, b) space where R²_indirect > 1 is possible. The color indicates the critical value of c' at the boundary. The gray area corresponds to the always bounded region."
#| fig-width: 7
#| fig-height: 6

eM_b <- 0.5
eY_b <- 0.5

boundary_grid <- expand.grid(
  a = seq(0.1, 3.0, length.out = 150),
  b = seq(0.1, 3.0, length.out = 150)
) |>
  dplyr::mutate(
    residuals     = b^2 * eM_b + eY_b,
    discriminant  = (a * b)^2 - residuals,
    viol_possible = discriminant > 0,
    cp_boundary   = dplyr::if_else(viol_possible,
                                   -a * b - sqrt(pmax(discriminant, 0)),
                                   NA_real_)
  )

ggplot(boundary_grid, aes(x = a, y = b)) +
  geom_tile(aes(fill = cp_boundary)) +
  geom_contour(aes(z = as.numeric(viol_possible)),
               breaks = 0.5, color = "black",
               linewidth = 1.5) +
  scale_fill_gradientn(
    colors  = c("steelblue", "white", "firebrick"),
    na.value = "grey88",
    name    = expression("critical c'")
  ) +
  annotate("text", x = 2.2, y = 2.3,
           label = "Violation\npossible",
           fontface = "bold", color = "white", size = 4) +
  annotate("text", x = 0.6, y = 0.5,
           label = "Always\nbounded",
           fontface = "bold", color = "grey40", size = 4) +
  labs(
    title    = "Boundary surface of the bounded region",
    subtitle = expression(sigma^2[eM] == 0.5 ~~ sigma^2[eY] == 0.5),
    x        = "Path coefficient a",
    y        = "Path coefficient b"
  ) -> p_boundary

print(p_boundary)
```

### 4.6 Summary of Results

The results of the empirical validation confirm the analytical implications of Theorem 1 and Corollary 1:

1.  **Corollary 1 is empirically confirmed**: in none of the `r format(nrow(results), big.mark=",")` tested cases is a violation observed with $|b^*| \leq 1$.

2.  $|b^*| > 1$ is a necessary condition for violation: 100% of the `r format(sum(results$exceeds), big.mark=",")` violation cases exhibit $|b^*| > 1$.

3.  **Within the realistic empirical range** ($|a^*|, |b^*| \leq 0.6$), corresponding to the vast majority of models in the social and behavioral sciences, **no violations are observed**.

4.  Violations are concentrated in extreme suppression scenarios where the standardized coefficients exceed unity — a condition that inherently signals structural instability of the estimated model.

Therefore, $R^2_{\text{indirect}}$ is **practically bounded** within $[0, 1]$ for all mediation models applied in the behavioral, social, and health sciences.

Having established this theoretical foundation and verified the mathematical boundaries of the index, the framework now transitions from analytical validation to computational implementation. The following section introduces the `r2_indirect` function, detailing its algorithmic structure and demonstrating its operationalization across single, parallel, and serial mediation architectures.

------------------------------------------------------------------------

### References

Wright, S. (1934). The method of path coefficients. *Annals of Mathematical Statistics*, *5*(3), 161–215.

------------------------------------------------------------------------

## 5. The $R^2_{\text{indirect}}$ Function (`r2_indirect`)

#FUNCTION FOR ESTIMATING SPECIFIC VARIANCE EXPLAINED BY INDIRECT EFFECTS

```{r}

r2_indirect <- function(model_fit, x_var, y_var, ..., n_boot = 500) {
  
  # -----------------------------------------------------------------
  # DEPENDENCY MANAGEMENT
  # Check if lavaan is loaded; if not, try to load it dynamically.
  # -----------------------------------------------------------------
  if (!"package:lavaan" %in% search()) {
    if (!require("lavaan", character.only = TRUE, quietly = TRUE)) {
      stop("Error: The 'lavaan' package is required but not installed. Please run install.packages('lavaan') first.")
    } else {
      message("Note: 'lavaan' package has been automatically loaded by r2_indirect().")
    }
  }
  
  # Capture the paths passed directly via ...
  paths_list <- list(...)
  num_paths <- length(paths_list)
  
  if (num_paths == 0) {
    stop("Error: Please specify at least one mediation path (e.g., path1 = c('a', 'b')).")
  }
  
  # Internal function for point estimates
  get_indices <- function(fit) {
    param <- parameterEstimates(fit)
    cov_matrix <- lavInspect(fit, "cov.ov")
    vX <- cov_matrix[x_var, x_var]
    vY <- cov_matrix[y_var, y_var]
    
    product_effects <- numeric(num_paths)
    
    for(j in 1:num_paths) {
      path_labels <- paths_list[[j]]
      # Extract and multiply the coefficients for the given path
      coef_values <- sapply(path_labels, function(l) param$est[param$label == l])
      product_effects[j] <- prod(coef_values)
    }
    
    # Calculate specific R2 indices
    r2_specific <- (product_effects)^2 * vX / vY
    
    # Calculate R2 for the TOTAL MEDIATED EFFECT
    r2_total <- (sum(product_effects))^2 * vX / vY
    
    return(c(r2_specific, r2_total))
  }
  
  # Point estimates on original data
  point_estimates <- get_indices(model_fit)
  
  # Bootstrapping procedure
  cat("Running Bootstrap on", num_paths, "specific path(s) + Total Mediated Effect...\n")
  orig_data <- as.data.frame(lavInspect(model_fit, "data"))
  boot_matrix <- matrix(NA, nrow = n_boot, ncol = num_paths + 1)
  
  pb <- txtProgressBar(min = 0, max = n_boot, style = 3)
  for(i in 1:n_boot) {
    boot_idx <- sample(1:nrow(orig_data), replace = TRUE)
    boot_data <- orig_data[boot_idx, ]
    fit_boot <- try(suppressWarnings(sem(model = model_fit@ParTable, data = boot_data)), silent = TRUE)
    
    if(!inherits(fit_boot, "try-error") && lavInspect(fit_boot, "converged")) {
      boot_matrix[i, ] <- get_indices(fit_boot)
    }
    setTxtProgressBar(pb, i)
  }
  close(pb)
  
  # Clean NA rows (non-converged bootstraps)
  boot_matrix <- na.omit(boot_matrix)
  
  # 95% Confidence Intervals
  ci_low <- apply(boot_matrix, 2, quantile, probs = 0.025)
  ci_high <- apply(boot_matrix, 2, quantile, probs = 0.975)
  
  # Format output names nicely based on user input
  path_names <- names(paths_list)
  if (is.null(path_names)) {
    path_names <- paste0("Path_", 1:num_paths, " [", sapply(paths_list, paste, collapse="*"), "]")
  } else {
    empty_names <- path_names == ""
    path_names[empty_names] <- paste0("Path_", which(empty_names))
    path_names <- paste0(path_names, " [", sapply(paths_list, paste, collapse="*"), "]")
  }
  
  effect_names <- c(path_names, "Total Mediated Effect")
  
  # Build final dataframe
  final_report <- data.frame(
    Effect = effect_names,
    r2_indirect = round(point_estimates, 5),
    CI_Low = round(ci_low, 5),
    CI_High = round(ci_high, 5),
    Perc = paste0(round(point_estimates * 100, 2), "%"),
    CI_Low_P = paste0(round(ci_low * 100, 2), "%"),
    CI_High_P = paste0(round(ci_high * 100, 2), "%")
  )
  
  return(final_report)
}
```

###EXAMPLE OF USAGE: results \<- r2_indirect(model_fit = lavaan.mod.fit, x_var = "X", y_var = "Y", ###User defined paths (Specify any type of indirect association as defined in the fitted lavaan model)### Single = c("path_a", "path_b"), Parallel1 = c("path_a1", "path_b1"), Parallel2 = c("path_a2", "path_b2"), Serial = c("path_a", "path_b", "path_d"), n_boot = 500)

print(results, row.names = FALSE)

### 5.1 Applyng the function to different mediation models

### 5.1.1 Single Mediator Models

```{r}

# ==============================================================================
# DATA GENERATION: SINGLE MEDIATOR
# ==============================================================================
simulate_data_single <- function(N, a, b, cp, seed = 123) {
  set.seed(seed)
  
  X  <- rnorm(N, mean = 0, sd = 1)
  
  # Mediator with residual variance = 1
  eM <- rnorm(N, mean = 0, sd = 1)
  M  <- a * X + eM
  
  # Outcome with residual variance = 1
  eY <- rnorm(N, mean = 0, sd = 1)
  Y  <- cp * X + b * M + eY
  
  return(data.frame(X = X, M = M, Y = Y))
}

# ---------------------------------------------------------------------
# Creation of the  Datasets
# ---------------------------------------------------------------------

# Dataset 1: N = 400 (Moderate positive effects)
df1 <- simulate_data_single(N = 400, a = 0.5, b = 0.5, cp = 0.4)

# Dataset 2: N = 500 (Strong positive effects)
df2 <- simulate_data_single(N = 500, a = 0.7, b = 0.6, cp = 0.3)

# Dataset 3: N = 600 (Mixed positive effects)
df3 <- simulate_data_single(N = 600, a = 0.4, b = 0.8, cp = 0.5)

# Dataset 4: N = 600 (TYPE 1 SUPPRESSION EFFECT: DIRECT NEGATIVE VS. INDIRECT POSITIVE)
# X depresses Y (negative cp), but the mediator increases Y
df4 <- simulate_data_single(N = 600, a = 0.6, b = 0.6, cp = -0.7)

# Dataset 4: N = 600 (TYPE 2 SUPPRESSION EFFECT: DIRECT POSITIVE VS. INDIRECT NEGATIVE)
# M depresses Y (negative ind effect), but the mediator increases Y
df5 <- simulate_data_single(N = 600, a = 0.6, b = -0.6, cp = 0.7)


# ==============================================================================
# LAVAAN MODEL DEFINITION
# ==============================================================================
# Load the required package
library(lavaan)

mod_simple <- "

# Paths
M ~ a * X
Y ~ cp * X + b * M

#Indirect effect
ind := a*b

"

fit1 <- sem(mod_simple, estimator = "MLR", data = df1)
fit2 <- sem(mod_simple, estimator = "MLR", data = df2)
fit3 <- sem(mod_simple, estimator = "MLR", data = df3)
fit4 <- sem(mod_simple, estimator = "MLR", data = df4)
fit5 <- sem(mod_simple, estimator = "MLR", data = df5)

summary(fit1, standardized = TRUE)
summary(fit2, standardized = TRUE)
summary(fit3, standardized = TRUE)
summary(fit4, standardized = TRUE)
summary(fit5, standardized = TRUE)


# ==============================================================================
# RUNNING r2_indirect ON RAW COEFFICIENTS (UNSTANDARDIZED b) SINGLE MEDIATOR
# ==============================================================================

cat("\n=== LAUNCHING r2_indirect ON RAW UNSTANDARDIZED DATA ===\n")
invisible(capture.output({
res_df1 <- r2_indirect(fit1, x_var = "X", y_var = "Y", Med_Path = c("a", "b"), n_boot = 200)
res_df2 <- r2_indirect(fit2, x_var = "X", y_var = "Y", Med_Path = c("a", "b"), n_boot = 200)
res_df3 <- r2_indirect(fit3, x_var = "X", y_var = "Y", Med_Path = c("a", "b"), n_boot = 200)
res_df4 <- r2_indirect(fit4, x_var = "X", y_var = "Y", Med_Path = c("a", "b"), n_boot = 200)
res_df5 <- r2_indirect(fit5, x_var = "X", y_var = "Y", Med_Path = c("a", "b"), n_boot = 200)
}))

cat("\n--> FUNCTION REPORT: DATASET 1 (N = 400)\n")
print(res_df1, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 2 (N = 500)\n")
print(res_df2, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 3 (N = 600)\n")
print(res_df3, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 4 (TYPE 1 SUPPRESSION, N = 600)\n")
print(res_df4, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 4 (TYPE 2 SUPPRESSION, N = 600)\n")
print(res_df5, row.names = FALSE)

# ==============================================================================
# METHODOLOGICAL CHECK: WRIGHT'S DECOMPOSITION FOR SINGLE MEDIATOR (5 BLOCKS)
# ==============================================================================

wright_proof_single <- function(model_fit, dataset_name = "Dataset") {
  
  # 1. Extract raw parameter estimates from the fitted model
  est <- parameterEstimates(model_fit)
  
  a_est  <- est$est[est$label == "a"]
  b_est  <- est$est[est$label == "b"]
  cp_est <- est$est[est$label == "cp"]
  
  # Extract residual variances (e_M and e_Y)
  var_eM <- est$est[est$op == "~~" & est$lhs == "M" & est$rhs == "M"]
  var_eY <- est$est[est$op == "~~" & est$lhs == "Y" & est$rhs == "Y"]
  
  # Extract observed variances from the sample covariance matrix (Fixed X and Y)
  cov_mat <- lavInspect(model_fit, "cov.ov")
  var_X   <- cov_mat["X", "X"]
  var_Y   <- cov_mat["Y", "Y"]
  
  # 2. Calculate the 5 components in RAW METRIC
  B1_Raw <- (cp_est)^2 * var_X
  B2_Raw <- (a_est * b_est)^2 * var_X
  B3_Raw <- 2 * cp_est * a_est * b_est * var_X
  B4_Raw <- (b_est)^2 * var_eM
  B5_Raw <- var_eY
  
  Total_Raw <- B1_Raw + B2_Raw + B3_Raw + B4_Raw + B5_Raw
  
  # 3. Calculate PROPORTIONS and PERCENTAGES
  P1 <- B1_Raw / var_Y; P2 <- B2_Raw / var_Y; P3 <- B3_Raw / var_Y
  P4 <- B4_Raw / var_Y; P5 <- B5_Raw / var_Y
  
  Total_Prop <- P1 + P2 + P3 + P4 + P5
  
  # --- PRINTING THE COMPREHENSIVE MATHEMATICAL CHECK REPORT ---
  cat("\n===========================================================================\n")
  cat(" WRIGHT DECOMPOSITION PROOF (SINGLE):", dataset_name, "\n")
  cat("===========================================================================\n")
  cat(sprintf(" Observed Variance of Y (Denominator) : %10.5f\n", var_Y))
  cat(sprintf(" Implied Variance of Y (Reconstructed): %10.5f\n", Total_Raw))
  cat(sprintf(" Discrepancy (Reconstruction Error)   : %10.5f\n", var_Y - Total_Raw))
  cat("---------------------------------------------------------------------------\n")
  cat(sprintf(" %-38s | %-10s | %-10s | %-10s\n", "COMPONENT", "RAW VAR", "PROPORTION", "PERCENT"))
  cat("---------------------------------------------------------------------------\n")
  cat(sprintf(" 1. Pure Direct Effect (cp^2)           | %10.5f | %10.5f | %8.2f %%\n", B1_Raw, P1, P1*100))
  cat(sprintf(" 2. Pure Indirect Effect ((ab)^2)       | %10.5f | %10.5f | %8.2f %%\n", B2_Raw, P2, P2*100))
  cat(sprintf(" 3. Wright Joint Cov (2*cp*a*b)         | %10.5f | %10.5f | %8.2f %%\n", B3_Raw, P3, P3*100))
  cat(sprintf(" 4. Independent M Variance (b^2*eM)     | %10.5f | %10.5f | %8.2f %%\n", B4_Raw, P4, P4*100))
  cat(sprintf(" 5. Y Residual/Error Variance (eY)      | %10.5f | %10.5f | %8.2f %%\n", B5_Raw, P5, P5*100))
  cat("---------------------------------------------------------------------------\n")
  cat(sprintf(" %-38s | %10.5f | %10.5f | %8.2f %%\n", "TOTAL CHECK", Total_Raw, Total_Prop, Total_Prop*100))
  cat("===========================================================================\n")
}

# ==============================================================================
# EXECUTE PROOF ON THE FITTED MODELS
# ==============================================================================
# Assuming models are already estimated in lavaan using raw data:
wright_proof_single(fit1, dataset_name = "Dataset 1 (N = 400)")
wright_proof_single(fit2, dataset_name = "Dataset 2 (N = 500)")
wright_proof_single(fit3, dataset_name = "Dataset 3 (N = 600)")
wright_proof_single(fit4, dataset_name = "Dataset 4 (TYPE 1 SUPPRESSION, N = 600)")
wright_proof_single(fit4, dataset_name = "Dataset 4 (TYPE 2 SUPPRESSION, N = 600)")
################################################################################
```

As demonstrated by the single-mediator outputs, the function accurately isolates the specific indirect variance while the accompanying algebraic proofs confirm a zero reconstruction error, even under opposing suppression conditions. Building upon this baseline architecture, the methodological framework may be reliably extended to accommodate simultaneous, competing pathways, as illustrated in the parallel mediation models below.

### 5.1.2 Models with Two Parallel Mediators

```{r}
# ==============================================================================
# DATA GENERATION: TWO PARALLEL MEDIATORS
# ==============================================================================

simulate_data_parallel <- function(N, a1, b1, a2, b2, cp, res_cov, seed = 123) {
  set.seed(seed)
  X <- rnorm(N, mean = 0, sd = 1)
  
  # Inject a shared unobserved variance component (creates residual covariance)
  shared_residual <- rnorm(N, mean = 0, sd = sqrt(res_cov))
  
  M1 <- a1 * X + shared_residual + rnorm(N, mean = 0, sd = sqrt(0.64 - res_cov))
  M2 <- a2 * X + shared_residual + rnorm(N, mean = 0, sd = sqrt(0.64 - res_cov))
  
  Y <- cp * X + b1 * M1 + b2 * M2 + rnorm(N, mean = 0, sd = 0.5)
  
  return(data.frame(X = X, M1 = M1, M2 = M2, Y = Y))
}

# ---------------------------------------------------------------------
# Creation of the  Datasets
# ---------------------------------------------------------------------

# Dataset 1: N = 600, Standard (All positive cooperative effects)
dfp1 <- simulate_data_parallel(N = 600, 
                               a1 = 0.4, b1 = 0.3, 
                               a2 = 0.3, b2 = 0.4, 
                               cp = 0.3, res_cov = 0.25, seed = 111)

# Dataset 2: N = 700, Standard (All positive cooperative effects)
dfp2 <- simulate_data_parallel(N = 700, 
                               a1 = 0.5, b1 = 0.25, 
                               a2 = 0.4, b2 = 0.3, 
                               cp = 0.35, res_cov = 0.20, seed = 222)

# Dataset 3: N = 700, Type 1 Suppression (Direct Positive vs. M1 (indirect) Negative)
dfp3 <- simulate_data_parallel(N = 700, 
                               a1 = 0.4, b1 = -0.4, 
                               a2 = 0.3, b2 = 0.3, 
                               cp = 0.4, res_cov = 0.30, seed = 333)

# Dataset 4: N = 700, Type 2 Suppression (Direct Positive vs. M1 & M2 (indirects) Negative)
dfp4 <- simulate_data_parallel(N = 700, 
                               a1 = 0.4, b1 = -0.4, 
                               a2 = -0.3, b2 = 0.4, 
                               cp = 0.5, res_cov = 0.28, seed = 444)

# Dataset 5: N = 700, Type 3 Suppression (Direct Negative vs. M1 & M2 (indirects) Positive)
dfp5 <- simulate_data_parallel(N = 700, 
                               a1 = 0.4, b1 = 0.4, 
                               a2 = 0.3, b2 = 0.4, 
                               cp = -0.5, res_cov = 0.28, seed = 444)

# ==============================================================================
# LAVAAN MODEL MODEL DEFINITION (df = 0)
# ==============================================================================
# Load the required package
library(lavaan)

parallel_model <- "
M1 ~ a1*X
M2 ~ a2*X
Y  ~ cp*X + b1*M1 + b2*M2
  
#Residual covariance for perfect identification (df=0)
  M1 ~~ M2
  
#Indirect effects
indM1 := a1*b1
indM2 := a2*b2
"

fitp1 <- sem(parallel_model, estimator = "MLR", data = dfp1)
fitp2 <- sem(parallel_model, estimator = "MLR", data = dfp2)
fitp3 <- sem(parallel_model, estimator = "MLR", data = dfp3)
fitp4 <- sem(parallel_model, estimator = "MLR", data = dfp4)
fitp5 <- sem(parallel_model, estimator = "MLR", data = dfp5)

summary(fitp1, standardized = TRUE)
summary(fitp2, standardized = TRUE)
summary(fitp3, standardized = TRUE)
summary(fitp4, standardized = TRUE)
summary(fitp5, standardized = TRUE)


# ==================================================================================
# RUNNING r2_indirect ON RAW COEFFICIENTS (UNSTANDARDIZED b) TWO PARALLEL MEDIATORS
# ==================================================================================

cat("\n=== LAUNCHING r2_indirect ON RAW UNSTANDARDIZED DATA ===\n")
invisible(capture.output({
res_dfp1 <- r2_indirect(fitp1, 
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        n_boot = 200)

res_dfp2 <- r2_indirect(fitp2, 
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        n_boot = 200)

res_dfp3 <- r2_indirect(fitp3,  
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        n_boot = 200)

res_dfp4 <- r2_indirect(fitp4, 
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        n_boot = 200)

res_dfp5 <- r2_indirect(fitp5, 
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        n_boot = 200)
}))

cat("\n--> FUNCTION REPORT: DATASET 1 (STANDARD, N = 600)\n")
print(res_dfp1, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 2 (STANDARD, N = 700)\n")
print(res_dfp2, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 3 (TYPE 1 SUPPRESSION, N = 700)\n")
print(res_dfp3, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 4 (TYPE 2 SUPPRESSION, N = 700)\n")
print(res_dfp4, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 5 (TYPE 3 SUPPRESSION, N = 700)\n")
print(res_dfp5, row.names = FALSE)


# ===============================================================================
# METHODOLOGICAL CHECK: WRIGHT'S DECOMPOSITION FOR PARALLEL MEDIATORS (10 BLOCKS)
# ===============================================================================
wright_proof_parallel <- function(model_fit, dataset_name = "Dataset") {
  
  est <- parameterEstimates(model_fit)
  
  # Regressions
  a1_est <- est$est[est$label == "a1"]
  b1_est <- est$est[est$label == "b1"]
  a2_est <- est$est[est$label == "a2"]
  b2_est <- est$est[est$label == "b2"]
  cp_est <- est$est[est$label == "cp"]
  
  # Residual Variances
  var_eM1 <- est$est[est$op == "~~" & est$lhs == "M1" & est$rhs == "M1"]
  var_eM2 <- est$est[est$op == "~~" & est$lhs == "M2" & est$rhs == "M2"]
  var_eY  <- est$est[est$op == "~~" & est$lhs == "Y"  & est$rhs == "Y"]
  
  # Residual Covariance of parallel mediators 
  cov_eM12_est <- est$est[est$op == "~~" & 
                            ((est$lhs == "M1" & est$rhs == "M2") | 
                               (est$lhs == "M2" & est$rhs == "M1"))]
  
  # Matrix Sample Variances
  cov_mat <- lavInspect(model_fit, "cov.ov")
  var_X   <- cov_mat["X", "X"]
  var_Y   <- cov_mat["Y", "Y"]
  
  # Compute the 10 Raw Components
  B1_Raw <- (cp_est)^2 * var_X
  B2_Raw <- (a1_est * b1_est)^2 * var_X
  B3_Raw <- (a2_est * b2_est)^2 * var_X
  B4_Raw <- 2 * cp_est * a1_est * b1_est * var_X
  B5_Raw <- 2 * cp_est * a2_est * b2_est * var_X
  B6_Raw <- 2 * (a1_est * b1_est) * (a2_est * b2_est) * var_X
  B7_Raw <- (b1_est)^2 * var_eM1
  B8_Raw <- (b2_est)^2 * var_eM2
  B9_Raw <- 2 * b1_est * b2_est * cov_eM12_est
  B10_Raw <- var_eY
  
  Total_Raw <- B1_Raw + B2_Raw + B3_Raw + B4_Raw + B5_Raw + B6_Raw + B7_Raw + B8_Raw + B9_Raw + B10_Raw
  
  # Proportions
  P1 <- B1_Raw / var_Y; P2 <- B2_Raw / var_Y; P3 <- B3_Raw / var_Y
  P4 <- B4_Raw / var_Y; P5 <- B5_Raw / var_Y; P6 <- B6_Raw / var_Y
  P7 <- B7_Raw / var_Y; P8 <- B8_Raw / var_Y; P9 <- B9_Raw / var_Y; P10 <- B10_Raw / var_Y
  
  Total_Prop <- P1 + P2 + P3 + P4 + P5 + P6 + P7 + P8 + P9 + P10
  
  # --- PRINTING THE COMPREHENSIVE MATHEMATICAL CHECK REPORT ---
  cat("\n===========================================================================\n")
  cat(" WRIGHT DECOMPOSITION PROOF (PARALLEL):", dataset_name, "\n")
  cat("===========================================================================\n")
  cat(sprintf(" Observed Variance of Y (Denominator) : %10.5f\n", var_Y))
  cat(sprintf(" Implied Variance of Y (Reconstructed): %10.5f\n", Total_Raw))
  cat(sprintf(" Discrepancy (Reconstruction Error)   : %10.5f\n", var_Y - Total_Raw))
  cat("---------------------------------------------------------------------------\n")
  cat(sprintf(" %-38s | %-10s | %-10s | %-10s\n", "COMPONENT", "RAW VAR", "PROPORTION", "PERCENT"))
  cat("---------------------------------------------------------------------------\n")
  cat(sprintf("  1. Pure Direct Effect (cp^2)          | %10.5f | %10.5f | %8.2f %%\n", B1_Raw, P1, P1*100))
  cat(sprintf("  2. Pure Indirect M1 ((a1*b1)^2)       | %10.5f | %10.5f | %8.2f %%\n", B2_Raw, P2, P2*100))
  cat(sprintf("  3. Pure Indirect M2 ((a2*b2)^2)       | %10.5f | %10.5f | %8.2f %%\n", B3_Raw, P3, P3*100))
  cat(sprintf("  4. Wright Joint Cov (2*cp*a1*b1)      | %10.5f | %10.5f | %8.2f %%\n", B4_Raw, P4, P4*100))
  cat(sprintf("  5. Wright Joint Cov (2*cp*a2*b2)      | %10.5f | %10.5f | %8.2f %%\n", B5_Raw, P5, P5*100))
  cat(sprintf("  6. Wright Joint Cov (2*a1*b1*a2*b2)   | %10.5f | %10.5f | %8.2f %%\n", B6_Raw, P6, P6*100))
  cat(sprintf("  7. Independent M1 Error (b1^2*eM1)    | %10.5f | %10.5f | %8.2f %%\n", B7_Raw, P7, P7*100))
  cat(sprintf("  8. Independent M2 Error (b2^2*eM2)    | %10.5f | %10.5f | %8.2f %%\n", B8_Raw, P8, P8*100))
  cat(sprintf("  9. Residual Joint Cov (2*b1*b2*cov)   | %10.5f | %10.5f | %8.2f %%\n", B9_Raw, P9, P9*100))
  cat(sprintf(" 10. Y Residual/Error Var (eY)          | %10.5f | %10.5f | %8.2f %%\n", B10_Raw, P10, P10*100))
  cat("---------------------------------------------------------------------------\n")
  cat(sprintf(" %-38s | %10.5f | %10.5f | %8.2f %%\n", "TOTAL CHECK", Total_Raw, Total_Prop, Total_Prop*100))
  cat("===========================================================================\n")
}

# ==============================================================================
# EXECUTE PROOF ON THE FITTED MODELS
# ==============================================================================
# Assuming models are already estimated in lavaan using raw data:
wright_proof_parallel(fitp1, "Dataset 1 (N=600, Standard)")
wright_proof_parallel(fitp2, "Dataset 2 (N=700, Standard)")
wright_proof_parallel(fitp3, "Dataset 3 (N=700, Type 1 Suppression)")
wright_proof_parallel(fitp4, "Dataset 4 (N=700, Type 2 Suppression)")
wright_proof_parallel(fitp5, "Dataset 5 (N=700, Type 3 Suppression)")
################################################################################

```

The evaluation of parallel models highlights the function's capacity to partition variance accurately across multiple indirect pathways while simultaneously accounting for the shared, unobserved residual covariances between mediators. This precise isolation of variance components is equally critical when mediators operate interdependently rather than concurrently, leading directly to the serial mediation structures examined next.

### 5.1.3 Models with Two Serial Mediators

```{r}
# ==============================================================================
# DATA GENERATION: TWO SERIAL MEDIATORS
# ==============================================================================

generate_serial_data <- function(N, a1, a2, d21, b1, b2, cp, seed = 17) {
  set.seed(seed)
  
  # Generating the independent variable X
  X <- rnorm(N, mean = 0, sd = 1)
  
  # Eq 1: M1 depends only on X
  e_M1 <- rnorm(N, mean = 0, sd = 1)
  M1 <- (a1 * X) + e_M1
  
  # Eq 2: M2 depends on X and M1 (Serial Path d21)
  e_M2 <- rnorm(N, mean = 0, sd = 1)
  M2 <- (a2 * X) + (d21 * M1) + e_M2
  
  # Eq 3: Y depends on X, M1 and M2
  e_Y <- rnorm(N, mean = 0, sd = 1)
  Y <- (cp * X) + (b1 * M1) + (b2 * M2) + e_Y
  
  return(data.frame(X, M1, M2, Y))
}

# ---------------------------------------------------------------------
# Creation of the  Datasets
# ---------------------------------------------------------------------

# Dataset 1: N = 600, Standard (All positive cooperative effects)
dfs1 <- generate_serial_data(N = 600, 
                             a1 = 0.5, a2 = 0.4, d21 = 0.5, 
                             b1 = 0.4, b2 = 0.5, cp = 0.3, 
                             seed = 101)  

# Dataset 2: N = 700, Standard (All positive, unbalanced weights on the serial path)
dfs2 <- generate_serial_data(N = 700, 
                             a1 = 0.6, a2 = 0.2, d21 = 0.6, 
                             b1 = 0.2, b2 = 0.5, cp = 0.4, 
                             seed = 102)        

# Dataset 3: N = 700, Type 1 Suppression (Direct Positive vs. M1 & Serial Negative)
dfs3 <- generate_serial_data(N = 700, 
                             a1 = 0.6, a2 = 0.2, d21 = 0.5, 
                             b1 = -0.5, b2 = 0.6, cp = 0.6, 
                             seed = 103)       

# Dataset 4: N = 700, Type 2 Suppression (Direct Positive vs. M2 & Serial Negative)
dfs4 <- generate_serial_data(N = 700, 
                             a1 = 0.6, a2 = 0.5, d21 = 0.4, 
                             b1 = 0.5, b2 = -0.6, cp = 0.2, 
                             seed = 104)       

# Dataset 5: N = 700, Type 3 Suppression (Direct, M1, M2, Positive vs. Serial Negative)
dfs5 <- generate_serial_data(N = 700, 
                             a1 = 0.6, a2 = 0.3, d21 = -0.7, 
                             b1 = 0.4, b2 = 0.5, cp = 0.5, 
                             seed = 105)      

# Dataset 5: N = 700, Type 4 Suppression (Direct, M1, M2, Positive vs. Serial Negative)
dfs6 <- generate_serial_data(N = 700, 
                             a1 = 0.6, a2 = 0.3, d21 = 0.7, 
                             b1 = 0.4, b2 = 0.5, cp = -0.5, 
                             seed = 105)  


# ==============================================================================
# LAVAAN MODEL MODEL DEFINITION (df = 0)
# ==============================================================================
# Load the required package
library(lavaan)

serial_model <- "
  M1 ~ a1*X
  M2 ~ a2*X + d*M1
  Y  ~ cp*X + b1*M1 + b2*M2

  #indirect effect via M1
  ind_M1 := a1 * b1
  
  #indirect effect via M2
  ind_M2 := a2 * b2
  
  #serial indirect effect via M1 and M2
  ind_serial := a1 * d * b2
"

fits1 <- sem(serial_model, estimator = "MLR", data = dfs1)
fits2 <- sem(serial_model, estimator = "MLR", data = dfs2)
fits3 <- sem(serial_model, estimator = "MLR", data = dfs3)
fits4 <- sem(serial_model, estimator = "MLR", data = dfs4)
fits5 <- sem(serial_model, estimator = "MLR", data = dfs5)
fits6 <- sem(serial_model, estimator = "MLR", data = dfs6)

summary(fits1, standardized = TRUE)
summary(fits2, standardized = TRUE)
summary(fits3, standardized = TRUE)
summary(fits4, standardized = TRUE)
summary(fits5, standardized = TRUE)
summary(fits6, standardized = TRUE)


# ================================================================================
# RUNNING r2_indirect ON RAW COEFFICIENTS (UNSTANDARDIZED b) TWO SERIAL MEDIATORS
# ================================================================================

cat("\n=== LAUNCHING r2_indirect ON RAW UNSTANDARDIZED DATA ===\n")
invisible(capture.output({
res_dfs1 <- r2_indirect(fits1, 
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        Serial_Path = c("a1", "d", "b2"),
                        n_boot = 200)

res_dfs2 <- r2_indirect(fits2, 
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        Serial_Path = c("a1", "d", "b2"),
                        n_boot = 200)

res_dfs3 <- r2_indirect(fits3,  
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        Serial_Path = c("a1", "d", "b2"),
                        n_boot = 200)

res_dfs4 <- r2_indirect(fits4, 
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        Serial_Path = c("a1", "d", "b2"),
                        n_boot = 200)

res_dfs5 <- r2_indirect(fits5, 
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        Serial_Path = c("a1", "d", "b2"),
                        n_boot = 200)

res_dfs6 <- r2_indirect(fits6, 
                        x_var = "X", 
                        y_var = "Y", 
                        Med_Path1 = c("a1", "b1"),
                        Med_Path2 = c("a2", "b2"),
                        Serial_Path = c("a1", "d", "b2"),
                        n_boot = 200)
}))

cat("\n--> FUNCTION REPORT: DATASET 1 (N = 600) STANDARD\n")
print(res_dfs1, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 2 (N = 700) STANDARD\n")
print(res_dfs2, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 3 (TYPE 1 SUPPRESSION, N = 700)\n")
print(res_dfs3, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 4 (TYPE 2 SUPPRESSION, N = 700)\n")
print(res_dfs4, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 5 (TYPE 3 SUPPRESSION, N = 700)\n")
print(res_dfs5, row.names = FALSE)

cat("\n--> FUNCTION REPORT: DATASET 6 (TYPE 4 SUPPRESSION, N = 700)\n")
print(res_dfs5, row.names = FALSE)


# ==============================================================================
# METHODOLOGICAL CHECK: WRIGHT'S DECOMPOSITION FOR SERIAL MEDIATORS (15 BLOCKS)
# ==============================================================================
wright_proof_serial <- function(model_fit, dataset_name = "Dataset") {
  
  est <- parameterEstimates(model_fit)
  
  # Structural paths
  a1_est <- est$est[est$label == "a1"]
  b1_est <- est$est[est$label == "b1"]
  a2_est <- est$est[est$label == "a2"]
  b2_est <- est$est[est$label == "b2"]
  d_est  <- est$est[est$label == "d"]
  cp_est <- est$est[est$label == "cp"]
  
  # Residual Variances
  var_eM1 <- est$est[est$op == "~~" & est$lhs == "M1" & est$rhs == "M1"]
  var_eM2 <- est$est[est$op == "~~" & est$lhs == "M2" & est$rhs == "M2"]
  var_eY  <- est$est[est$op == "~~" & est$lhs == "Y"  & est$rhs == "Y"]
  
  # Matrix Sample Variances
  cov_mat <- lavInspect(model_fit, "cov.ov")
  var_X   <- cov_mat["X", "X"]
  var_Y   <- cov_mat["Y", "Y"]
  
  # Compute the 15 Raw Components
  # --- 1. Effects originating from X ---
  B1_Raw  <- (cp_est)^2 * var_X
  B2_Raw  <- (a1_est * b1_est)^2 * var_X
  B3_Raw  <- (a2_est * b2_est)^2 * var_X
  B4_Raw  <- (a1_est * d_est * b2_est)^2 * var_X
  B5_Raw  <- 2 * cp_est * (a1_est * b1_est) * var_X
  B6_Raw  <- 2 * cp_est * (a2_est * b2_est) * var_X
  B7_Raw  <- 2 * cp_est * (a1_est * d_est * b2_est) * var_X
  B8_Raw  <- 2 * (a1_est * b1_est) * (a2_est * b2_est) * var_X
  B9_Raw  <- 2 * (a1_est * b1_est) * (a1_est * d_est * b2_est) * var_X
  B10_Raw <- 2 * (a2_est * b2_est) * (a1_est * d_est * b2_est) * var_X
  
  # --- 2. Effects originating from eM1 
  B11_Raw <- (b1_est)^2 * var_eM1
  B12_Raw <- (d_est * b2_est)^2 * var_eM1
  B13_Raw <- 2 * b1_est * (d_est * b2_est) * var_eM1
  
  # --- 3. Effects originating from eM2 and eY ---
  B14_Raw <- (b2_est)^2 * var_eM2
  B15_Raw <- var_eY
  
  Total_Raw <- sum(B1_Raw, B2_Raw, B3_Raw, B4_Raw, B5_Raw, B6_Raw, B7_Raw, 
                   B8_Raw, B9_Raw, B10_Raw, B11_Raw, B12_Raw, B13_Raw, B14_Raw, B15_Raw)
  
  # Proportions
  P1  <- B1_Raw / var_Y;  P2  <- B2_Raw / var_Y;  P3  <- B3_Raw / var_Y
  P4  <- B4_Raw / var_Y;  P5  <- B5_Raw / var_Y;  P6  <- B6_Raw / var_Y
  P7  <- B7_Raw / var_Y;  P8  <- B8_Raw / var_Y;  P9  <- B9_Raw / var_Y
  P10 <- B10_Raw / var_Y; P11 <- B11_Raw / var_Y; P12 <- B12_Raw / var_Y
  P13 <- B13_Raw / var_Y; P14 <- B14_Raw / var_Y; P15 <- B15_Raw / var_Y
  
  Total_Prop <- sum(P1, P2, P3, P4, P5, P6, P7, P8, P9, P10, P11, P12, P13, P14, P15)
  
  # --- PRINTING THE COMPREHENSIVE MATHEMATICAL CHECK REPORT ---
  cat("\n===========================================================================\n")
  cat(" WRIGHT DECOMPOSITION PROOF (SERIAL):", dataset_name, "\n")
  cat("===========================================================================\n")
  cat(sprintf(" Observed Variance of Y (Denominator) : %10.5f\n", var_Y))
  cat(sprintf(" Implied Variance of Y (Reconstructed): %10.5f\n", Total_Raw))
  cat(sprintf(" Discrepancy (Reconstruction Error)   : %10.5f\n", var_Y - Total_Raw))
  cat("---------------------------------------------------------------------------\n")
  cat(sprintf(" %-38s | %-10s | %-10s | %-10s\n", "COMPONENT", "RAW VAR", "PROPORTION", "PERCENT"))
  cat("---------------------------------------------------------------------------\n")
  cat(sprintf("  1. Pure Direct Effect (cp^2)          | %10.5f | %10.5f | %8.2f %%\n", B1_Raw,  P1,  P1*100))
  cat(sprintf("  2. Pure Indirect M1 ((a1*b1)^2)       | %10.5f | %10.5f | %8.2f %%\n", B2_Raw,  P2,  P2*100))
  cat(sprintf("  3. Pure Indirect M2 ((a2*b2)^2)       | %10.5f | %10.5f | %8.2f %%\n", B3_Raw,  P3,  P3*100))
  cat(sprintf("  4. Pure Serial ((a1*d*b2)^2)          | %10.5f | %10.5f | %8.2f %%\n", B4_Raw,  P4,  P4*100))
  cat(sprintf("  5. Wright Joint Cov (2*cp*a1*b1)      | %10.5f | %10.5f | %8.2f %%\n", B5_Raw,  P5,  P5*100))
  cat(sprintf("  6. Wright Joint Cov (2*cp*a2*b2)      | %10.5f | %10.5f | %8.2f %%\n", B6_Raw,  P6,  P6*100))
  cat(sprintf("  7. Wright Joint Cov (2*cp*a1*d*b2)    | %10.5f | %10.5f | %8.2f %%\n", B7_Raw,  P7,  P7*100))
  cat(sprintf("  8. Wright Joint Cov (2*a1*b1*a2*b2)   | %10.5f | %10.5f | %8.2f %%\n", B8_Raw,  P8,  P8*100))
  cat(sprintf("  9. Wright Joint Cov (2*a1*b1*a1*d*b2) | %10.5f | %10.5f | %8.2f %%\n", B9_Raw,  P9,  P9*100))
  cat(sprintf(" 10. Wright Joint Cov (2*a2*b2*a1*d*b2) | %10.5f | %10.5f | %8.2f %%\n", B10_Raw, P10, P10*100))
  cat(sprintf(" 11. Indep. M1 Error via b1 (b1^2*eM1)  | %10.5f | %10.5f | %8.2f %%\n", B11_Raw, P11, P11*100))
  cat(sprintf(" 12. Indep. M1 Error via M2 ((d*b2)^2*e)| %10.5f | %10.5f | %8.2f %%\n", B12_Raw, P12, P12*100))
  cat(sprintf(" 13. M1 Error Joint Cov (2*b1*d*b2*eM1) | %10.5f | %10.5f | %8.2f %%\n", B13_Raw, P13, P13*100))
  cat(sprintf(" 14. Indep. M2 Error (b2^2*eM2)         | %10.5f | %10.5f | %8.2f %%\n", B14_Raw, P14, P14*100))
  cat(sprintf(" 15. Y Residual/Error Var (eY)          | %10.5f | %10.5f | %8.2f %%\n", B15_Raw, P15, P15*100))
  cat("---------------------------------------------------------------------------\n")
  cat(sprintf(" %-38s | %10.5f | %10.5f | %8.2f %%\n", "TOTAL CHECK", Total_Raw, Total_Prop, Total_Prop*100))
  cat("===========================================================================\n")
}

# ==============================================================================
# EXECUTE PROOF ON THE FITTED MODELS
# ==============================================================================
# Assuming models are already estimated in lavaan using raw data:
wright_proof_serial(fits1, "Dataset 1 (N=600, Standard)")
wright_proof_serial(fits2, "Dataset 2 (N=700, Standard)")
wright_proof_serial(fits3, "Dataset 3 (N=700, Type 1 Suppression)")
wright_proof_serial(fits4, "Dataset 4 (N=700, Type 2 Suppression)")
wright_proof_serial(fits5, "Dataset 3 (N=700, Type 3 Suppression)")
wright_proof_serial(fits6, "Dataset 4 (N=700, Type 4 Suppression)")
################################################################################
```

The successful application of the 15-block decomposition to serial structures validates the metric's scalability across multi-stage causal chains (i.e., serial mediation). Having established a mathematically sound effect size index capable of handling single, parallel, and serial mediation architectures, the framework may now address the inferential requirements of prospective sample size planning and sensitivity analysis.

## 6. A-Priori Power Analysis and Sensitivity Analysis based on $R^2_{\text{indirect}}$

### 6.1 Theoretical Framework: Joint Significance Power

Statistical power in mediation analysis has been evaluated using individual path coefficients (or correlation) rather than a consolidated effect size metric. One of This framework operationalizes power analysis directly through the **Joint Significance Criterion**. Under this approach, an indirect effect is deemed statistically significant if and only if every single constituent path coefficient within the structural chain ($X \rightarrow M \rightarrow Y$) is significantly different from zero at a designated alpha level ($\alpha$).

Assuming the sampling distributions of the path coefficients are asymptotically independent, the joint power of the indirect pathway ($\text{Power}_{\text{joint}}$) is defined as the product of the individual statistical powers of its $k$ constituent paths:

$$\text{Power}_{\text{joint}} = \prod_{i=1}^{k} \text{Power}_i$$

Because $\text{Power}_i \le 1$, the joint power decreases as the length of the mediation chain ($k\_paths$) increases. This mathematical reality requires sample size planning to account for the compounding erosion of power across multi-stage structures. Since joint power relies on the multiplicative probability of multiple independent statistical tests, the exact same aggregate effect size may emerge from theoretically infinite configurations of path coefficients. To resolve this parameter indeterminacy for reliable sample size planning, it is necessary to establish mathematical boundary conditions, as detailed in the subsequent section.

### 6.2 Mathematical Boundary Profiles: Balanced vs. Bottleneck Scenarios

A known mathematical property of any product-based effect size — including the unstandardized product $a \cdot b$, the standardized product $\beta_a \cdot \beta_b$, and the variance-accounted-for metric $R^2_{\text{indirect}}$ — is that an infinite number of path combinations may yield the exact same global effect size. For instance, an $R^2_{\text{indirect}} = 0.01$ may result from two moderate paths or from one extremely large path paired with a near-zero path.

To resolve this indeterminacy without relying on arbitrary parameter assumptions, this framework implements a boundary-profile method that defines two structural extremes:

- **The Balanced Scenario (Best Case):** This profile assumes that the underlying structural mechanism is perfectly symmetrical. The total variance explained by the indirect effect is distributed evenly across all $k$ paths. For a target indirect coefficient $K = \sqrt{R^2_{\text{indirect}}}$, the standardized coefficient for each path is calculated as: $$\beta_{\text{balanced}} = K^{\frac{1}{k}}$$ Because statistical power is a non-linear function maximized when the minimum effect size in a chain is as large as possible, the balanced scenario yields the highest possible joint power for a given $R^2_{\text{indirect}}$. It provides the lower bound for the required sample size ($N$).

- **The Bottleneck Scenario (Worst Case):** This profile assumes extreme structural asymmetry. It does not necessarly represent a realistic empirical model, but rather a mathematical "stress test" to find the absolute lower bound of the required effect. One path acts as a severe statistical restriction (the "bottleneck", or implied minimum beta), while the remaining $k-1$ paths are artificially maximized up to a user-defined empirical ceiling (`max_beta`, default to 0.80). The bottleneck path coefficient ($\beta_{\text{min}}$) is formulated as: $$\beta_{\text{min}} = \frac{K}{(\beta_{\text{max}})^{k-1}}$$ By forcing all other paths to their theoretical limit (where a coefficient of 0.80 would border on tautology in social sciences), this scenario identifies the minimum possible strength required from the weakest link in the chain to still achieve the target power. By calculating both the balanced maximum and the bottleneck minimum, these profiles establish a definitive bounding range for sample size requirements without relying on arbitrary parameter guessing. However, precise power estimation also requires dynamic accounting of residual variances and structural penalties, which are integrated into the mathematical engine as follows.

### 6.3 Error Variance Adjustments and Multi-Mediator Penalties

The power engine does not treat the outcome variable's residual variance as a static parameter. Following Wright’s path tracing rules, the residual variance of the outcome ($\sigma^2_{eY}$) is dynamically computed by subtracting all explained components from the total unit variance ($1$):

$$\sigma^2_{eY} = 1 - \left(c'^2 + K^2 + 2c'K\right) - R^2_{\text{parallel}} - R^2_{\text{covariates}}$$

Where $c'$ is the direct effect, $K$ is the indirect path coefficient ($\sqrt{R^2_{\text{indirect}}}$), $2c'K$ is the joint covariance block, $R^2_{\text{parallel}}$ is the variance explained by alternative parallel paths, and $R^2_{\text{covariates}}$ represents external control variables.

Furthermore, when evaluating a target mediator embedded within a parallel mediation design, the framework applies an analytic Variance Inflation Factor (VIF) penalty to account for collinearity among mediators:

$$\text{VIF} = \frac{1}{1 - R^2_{\text{m\_overlap}}}$$

This penalty inflates the standard errors of the path coefficients, capturing the loss of statistical precision that occurs when parallel mediators share overlapping variance ($R^2_{\text{m\_overlap}}$). Incorporating these dynamic variance adjustments ensures that standard errors reflect the true statistical environment of the targeted pathway, appropriately penalizing for mediator collinearity. To prevent the "bottelneck" scenarios from pushing theoretical path coefficients into mathematically possible but empirically absurd ranges, an absolute structural limit may be enforced.

### 6.4 The Empirical Guardrail (`max_beta`)

In structural equation modeling, standardized path coefficients represent partial correlations or standardized regression weights. Mathematically, they are generally bounded by an absolute value of $1.0$. Furthermore, in empirical behavioral and social sciences, it is extremely rare for a single structural path to exceed $0.80$ without indicating severe theoretical redundancy or multicollinearity issues.

When the algorithm computes the Bottleneck Scenario, it forces one path to be as large as possible to evaluate the worst-case structural asymmetry. Without a cap, the mathematical engine could theoretically maximize this path up to $0.99$ to satisfy the required $R^2_{\text{indirect}}$, resulting in an empirically absurd profile.

To prevent this, the framework introduces the `max_beta` parameter as an empirical guardrail (default to $0.80$). It strictly restricts the maximum possible magnitude for any single constituent path in the mediation chain. If achieving a specific effect size requires a structural path that exceeds this limit, the algorithm halts and correctly flags the scenario as `"Unreachable"`, acknowledging that such a structure violates empirical plausibility.

The enforcement of this empirical guardrail preserves the ecological validity of the power analysis, guaranteeing that simulated bottleneck scenarios remain within the bounds of plausible behavioral and social science research. With these theoretical constraints defined, the framework translates these principles into an operational algorithm.

------------------------------------------------------------------------

## 7. Power Analysis for Single, Parallel, and Serial Mediation (`wright_power`)

### 7.1 Algorithmic Mechanics

The `wright_power` function computes the required sample size ($N$) to detect a target $R^2_{\text{indirect}}$ by executing a deterministic grid-search optimization across a range of $N \in [10, 5000]$.

For each iteration of $N$, the function executes the following mathematical sequence:

1.  **Standard Error Calculation:** It calculates the standard error for any given path coefficient ($\beta$) based on the current sample size $N$ and the dynamic residual variance: $$\text{SE}_{\beta} = \sqrt{\frac{\sigma^2_{eY}}{N \cdot (1 - \beta^2)}} \cdot \sqrt{\text{VIF}}$$
2.  **Individual Path Power:** It computes the statistical power for each distinct structural pathway using the standard normal cumulative distribution function ($\Phi$): $$\text{Power}_{\beta} = \Phi\left(\frac{\beta}{\text{SE}_{\beta}} - z_{1 - \frac{\alpha}{2}}\right)$$
3.  **Joint Power Aggregation:** It aggregates the individual path powers into the final joint power metric using a multiplicative probability function:
    - **For the Balanced Scenario:** $\text{Power}_{\text{joint}} = (\text{Power}_{\text{balanced}})^k$
    - **For the Bottleneck Scenario:** $\text{Power}_{\text{joint}} = \text{Power}_{\text{min}} \cdot (\text{Power}_{\text{max}})^{k-1}$
4.  **Identifying Required Sample Size:** The sequence iterates, increasing $N$, until it identifies the exact sample sizes where the calculated joint power successfully crosses the user-specified threshold (`power_target`).This deterministic grid-search ensures absolute computational precision in identifying the minimal $N$ required to achieve the desired joint statistical power. The exact parametrization of this search, however, fundamentally shifts depending on the targeted structural configuration, requiring specific architectural adaptations based on the hypothesized mediation model.

### 7.2 Structural Adaptation (Single, Parallel, and Serial Models)

The function adapts to different structural architectures via specific execution arguments:

- **Single Mediation:** Evaluated by setting `k_paths = 2`, `r2_parallel = 0`, and `r2_m_overlap = 0`.
- **Parallel Mediation:** Incorporates the variance drainage of concurrent pathways via `r2_parallel` and handles the structural collinearity of competing mediators by setting `r2_m_overlap > 0` (triggering the VIF penalty).
- **Serial Mediation:** Evaluated by scaling `k_paths` to $3$ or more structural steps. The mathematical engine automatically projects the compounding power degradation across the multi-stage chain.

Additionally, the algorithm also accounts for the directionality of the structural paths via the `type` parameter. Because the decomposition of variance includes a joint covariance block ($2c'K$), the signs of the effects dictate the final residual error of the outcome ($\sigma^2_{eY}$):

- **Consistent Models (`type = "consistent"`):** The direct effect ($c'$) and the indirect effect have the same sign. They cooperate to explain variance in $Y$, mathematically shrinking the residual error. This deflation of error variance tightens the standard errors of the paths, thereby increasing statistical power and lowering the required sample size.
- **Suppression Models (`type = "suppression"`):** The direct and indirect effects have opposite signs. Their mathematical combination cancels out explained variance, artificially inflating the residual error ($\sigma^2_{eY}$). This inflation bloats the standard errors, reducing power and requiring a larger sample size to detect the same $R^2_{\text{indirect}}$.
- **Explore Mode (`type = "explore"`):** The default setting, which forces the engine to run the optimization twice — once under a consistent regime and once under a suppression regime. This provides the researcher with a complete bounding profile, defining both the best-case and worst-case sample size requirements based on path directionality.

```{r}
#' Wright Mediation Power Analysis (A-Priori)
#' 
#' Computes the required Sample Size (N) to detect a specific indirect R2,
#' returning the Best (Balanced) and Worst (Bottleneck) scenarios.
#' 
#' @param r2_indirect Target variance of Y explained by the indirect effect.
#' @param type "explore" (default), "consistent", or "suppression".
#' @param k_paths Number of multiplicative paths in the indirect block (default = 2).
#' @param c_prime Expected direct effect coefficient (default = 0.20).
#' @param r2_m_overlap: 
#' Variance of the target mediator explained by other parallel mediators (VIF penalty).
#' Tip: For 2 mediators with a correlation of 0.30, enter the squared value (0.30^2 = 0.09).
#' Tip: For 3+ mediators, enter the multiple R2 of the target regressed on the others.
#' @param r2_parallel Variance of Y explained by other parallel mediators.
#' @param r2_covariates Variance of Y explained by external control variables.
#' @param power_target Desired joint statistical power threshold (default = 0.80).
#' @param alpha Significance level (default = 0.05).
#' @param max_beta Maximum allowed single standardized coefficient (guardrail, default = 0.80).

wright_power <- function(r2_indirect = NULL, target_beta_path = NULL, type = "explore", 
                         k_paths = 2, c_prime = 0.20,
                         r2_m_overlap = 0, r2_parallel = 0, r2_covariates = 0,
                         power_target = 0.80, alpha = 0.05, max_beta = 0.80) {
  
  # 1. INPUT VALIDATION & ARCHITECTURAL GUARDRAIL
  if(is.null(r2_indirect) && is.null(target_beta_path)) {
    stop("You must provide either 'r2_indirect' or 'target_beta_path'.")
  }
  if(!is.null(r2_indirect) && !is.null(target_beta_path)) {
    stop("Please provide either 'r2_indirect' or 'target_beta_path', not both.")
  }
  
  if(!is.null(r2_indirect) && (r2_indirect <= 0 | r2_indirect >= 1)) stop("r2_indirect must be between 0 and 1.")
  if(!is.null(target_beta_path) && (target_beta_path <= 0 | target_beta_path >= 1)) stop("target_beta_path must be between 0 and 1.")
  
  if(r2_covariates < 0 | r2_covariates >= 0.90) stop("r2_covariates must be between 0 and 0.90.")
  if(r2_parallel < 0 | r2_parallel >= 0.90) stop("r2_parallel must be between 0 and 0.90.")
  if(r2_m_overlap < 0 | r2_m_overlap >= 1) stop("r2_m_overlap must be strictly between 0 and 1.")
  if(k_paths < 2) stop("k_paths must be at least 2.")
  
  if(k_paths >= 4) {
    warning("Models with k >= 4 (3 or more serial mediators) are highly susceptible to severe variance attenuation and measurement error accumulation. Results are provided for mathematical evaluation, but empirical implementation requires extreme caution.", call. = FALSE)
  }
  
  # 2. TRANSLATE TARGET BETA TO R2 (If necessary)
  if(!is.null(target_beta_path)) {
    r2_indirect <- (target_beta_path^k_paths)^2
  }
  
  # 3. CALCULATE IMPLIED BETAS & EFFECT GUARDRAIL
  K_val <- sqrt(r2_indirect)
  beta_bal_implied <- K_val^(1/k_paths)
  beta_bot_implied <- K_val / (max_beta^(k_paths - 1))
  
  if(beta_bal_implied > 0.50 || (beta_bot_implied > 0.50 && beta_bot_implied <= max_beta)) {
    warning("The requested effect size implies a local step effect (beta) > 0.50. This is unusually large for social sciences.", call. = FALSE)
  }
  
  # 4. PARALLEL MEDIATION PENALTY (Analytic VIF)
  vif_penalty <- 1 / (1 - r2_m_overlap)
  
  # 5. CORE POWER ENGINE
  get_n <- function(cp) {
    K <- sqrt(r2_indirect)
    min_beta <- K / (max_beta^(k_paths - 1))
    
    # Skip if constraints violate the geometric guardrail
    if(min_beta > max_beta) return(c(NA, NA))
    
    # Compute residual variance of Y (incorporating covariates and parallel mediators)
    var_error_y <- 1 - (cp^2 + K^2 + 2 * cp * K) - r2_covariates - r2_parallel
    if(var_error_y <= 0) var_error_y <- 0.01
    
    # Grid search for N
    n_grid <- seq(10, 5000, by = 1)
    n_balanced <- NA
    n_bottleneck <- NA
    
    for(n in n_grid) {
      # --- BALANCED SCENARIO ---
      beta_bal <- K^(1/k_paths)
      se_bal <- sqrt(var_error_y / (n * (1 - beta_bal^2))) * sqrt(vif_penalty)
      pow_bal <- pnorm((beta_bal / se_bal) - qnorm(1 - alpha/2))
      joint_pow_bal <- pow_bal^k_paths
      
      # --- BOTTLENECK SCENARIO ---
      se_min <- sqrt(var_error_y / (n * (1 - min_beta^2))) * sqrt(vif_penalty)
      se_max <- sqrt(var_error_y / (n * (1 - max_beta^2))) * sqrt(vif_penalty)
      pow_min <- pnorm((min_beta / se_min) - qnorm(1 - alpha/2))
      pow_max <- pnorm((max_beta / se_max) - qnorm(1 - alpha/2))
      joint_pow_bot <- pow_min * (pow_max^(k_paths - 1))
      
      # Capture threshold crossing points
      if(is.na(n_balanced) && joint_pow_bal >= power_target) n_balanced <- n
      if(is.na(n_bottleneck) && joint_pow_bot >= power_target) n_bottleneck <- n
      
      # Break optimization if both profiles are populated
      if(!is.na(n_balanced) && !is.na(n_bottleneck)) break
    }
    return(c(n_balanced, n_bottleneck))
  }
  
  # 6. OUTPUT GENERATION
  cat("\n==================================================================================\n")
  cat("                WRIGHT MEDIATION A PRIORI POWER ANALYSIS                \n")
  cat("==================================================================================\n")
  if(!is.null(target_beta_path)) {
    cat(sprintf("Target Path Beta:      %.3f (Translates to R2 indirect: %.4f)\n", target_beta_path, r2_indirect))
  } else {
    cat(sprintf("Target Indirect R2:    %.4f \n", r2_indirect))
  }
  cat(sprintf("Structural Paths:      %d \n", k_paths))
  cat(sprintf("Direct Effect (c'):    %.2f (Base Magnitude)\n", abs(c_prime)))
  cat(sprintf("Other Mediators (R2):  %.2f (Variance in Y)\n", r2_parallel))
  cat(sprintf("Target M Overlap (R2): %.2f (Collinearity VIF = %.2f)\n", r2_m_overlap, vif_penalty))
  cat(sprintf("Covariates (R2):       %.2f \n", r2_covariates))
  cat(sprintf("Target Joint Power:    %.2f \n", power_target))
  cat("----------------------------------------------------------------------------------\n")
  cat(sprintf("%-12s %-12s %-18s %-18s %-12s\n", "SCENARIO", "MODEL TYPE", "IMPLIED MIN BETA", "IMPLIED MAX BETA", "REQUIRED N"))
  cat("----------------------------------------------------------------------------------\n")
  
  scenarios <- if(type == "explore") c("consistent", "suppression") else type
  
  for(t in scenarios) {
    cp_val <- if(t == "consistent") abs(c_prime) else -abs(c_prime)
    t_label <- if(t == "consistent") "Consistent" else "Suppression"
    
    results <- get_n(cp_val)
    
    n_bal <- if(!is.na(results[1])) sprintf("%d", results[1]) else "Unreachable"
    n_bot <- if(!is.na(results[2])) sprintf("%d", results[2]) else "Unreachable"
    
    b_bal_min_str <- sprintf("%.3f", beta_bal_implied)
    b_bal_max_str <- sprintf("%.3f", beta_bal_implied)
    
    b_bot_min_str <- sprintf("%.3f", beta_bot_implied)
    b_bot_max_str <- sprintf("%.3f (x%d)", max_beta, k_paths - 1)
    
    cat(sprintf("%-12s %-12s %-18s %-18s %-12s\n", "Balanced", t_label, b_bal_min_str, b_bal_max_str, n_bal))
    cat(sprintf("%-12s %-12s %-18s %-18s %-12s\n", "Bottleneck", t_label, b_bot_min_str, b_bot_max_str, n_bot))
    if(t != scenarios[length(scenarios)]) cat("- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -\n")
  }
  cat("==================================================================================\n")
  cat("Method: Joint Significance Power.\n\n")
}


################################################################################
###Function usage example!

#Single mediator model
wright_power(r2_indirect = 0.0081,
             type = "explore",
             k_paths = 2,
             c_prime = 0.30,
             r2_m_overlap = 0, 
             r2_parallel = 0, 
             r2_covariates = 0,
             power_target = 0.80,
             alpha = 0.05,
             max_beta = 0.80)

#Two parallel mediators model
wright_power(r2_indirect = 0.0081, 
             type = "explore", 
             k_paths = 2,
             c_prime = 0.30,
             r2_m_overlap = 0.09,
             r2_parallel = 0.09,
             r2_covariates = 0,
             power_target = 0.80, 
             alpha = 0.05,
             max_beta = 0.80)

#Two serial mediators model
wright_power(r2_indirect = 0.000729,
             type = "explore",
             k_paths = 3,
             r2_m_overlap = 0, 
             r2_parallel = 0, 
             r2_covariates = 0,
             power_target = 0.80,
             alpha = 0.05,
             max_beta = 0.80)

```

The execution of these models shows the algorithm's flexibility in managing single, parallel, and serial configurations of a mediation model while meticulously accounting for the statistical impact of consistent versus suppression dynamics. Thus, integrating the exact algebraic decomposition of $R^2_{\text{indirect}}$ with this boundary-profile power analysis provides researchers with a comprehensive, mathematically validated toolkit for standardizing effect sizes and rigorous sample size planning in structural mediation.

------------------------------------------------------------------------

## 8. Sensitivity Analysis for the MDES (`wright_sensitivity`)

### 8.1 Algorithmic Mechanics

When the sample size ($N$) is predetermined by resource constraints, retrospective design limitations, or existing datasets, the analytical objective shifts from finding $N$ to defining the **Minimum Detectable Effect Size (MDES)**. The `wright_sensitivity` function executes this by reverse-engineering the power calculation through a step-by-step search procedure:

1.  **The Grid Search:** The function initiates a loop, starting from a near-zero indirect effect ($R^2_{\text{indirect}} = 0.0001$) and incrementally increasing it step-by-step up to ($R^2_{\text{indirect}} = 0.8000$).
2.  **Parameter Extraction:** For each discrete variance value in the grid, the algorithm calculates exactly what the underlying path coefficients ($\beta$) would need to be under both the Balanced and Bottleneck structural profiles.
3.  **Error Variance Computation:** Using Wright’s decomposition, it calculates the dynamic residual variance of the outcome ($\sigma^2_{eY}$) associated with that specific structural configuration, accounting for any suppression or consistency effects.
4.  **Power Evaluation:** Using the fixed sample size $N$ provided by the user, it calculates the joint statistical power that would be achieved for that specific $R^2_{\text{indirect}}$.
5.  **Identifying the MDES:** The sequence iterates until the exact moment the calculated joint power crosses the user's specified threshold (e.g., $0.80$). The loop then stops, and that specific $R^2_{\text{indirect}}$ value is returned as the MDES. It represents the absolute smallest variance proportion the study is statistically equipped to detect reliably. By iteratively reversing the power logic, the algorithm shifts the analytical focus from prospective planning to retrospective evaluation. This ensures that even under strict empirical or resource constraints, researchers may define mathematically boundaries for their theoretical expectations.

### 8.2 Structural Interpretation of Outputs

The output of `wright_sensitivity` provides a clear assessment of a study's empirical threshold by deconstructing the required effect into specific path coefficients:

| Scenario | Implied MIN Beta | Implied MAX Beta | Required R2 (MDES) | Interpretation |
|:---|:---|:---|:---|:---|
| **Balanced** | *Calculated* | *Identical to MIN* | *Value* | The absolute lowest variance proportion the study can detect under optimal, symmetrical path conditions. All paths share the exact same required coefficient. |
| **Bottleneck** | *Calculated* | *Fixed (e.g., 0.80)* | *Value* | The minimum variance proportion required if the model relies on a highly asymmetric, weak pathway. The MAX Beta shows the artificial ceiling applied to $k-1$ paths, while the MIN Beta shows the required strength of the weakest link. |

**Understanding MIN vs. MAX Beta:** The inclusion of the "Implied MIN Beta" and "Implied MAX Beta" columns serves to ground the global target parameter ($R^2$) in local, empirical reality. \* In the **Balanced** scenario, MIN and MAX betas are identical, showing the required coefficient if the effect is perfectly distributed. \* In the **Bottleneck** scenario, the MAX beta is deliberately pushed to the limit (e.g., 0.80) to conduct a mathematical stress test. Researchers should evaluate the Implied MIN Beta: if achieving the required $R^2$ demands a MIN Beta that is implausibly large (e.g., $> 0.50$) even when all other paths are operating at absolute theoretical maximums, the study is severely underpowered for practical applications.

If a calculated MDES scenario returns `"Unreachable"`, it signifies that the structural parameters, background covariates, or collinearity penalties drain so much variance from the system that it is mathematically impossible to achieve the target power at that sample size.

```{r}
#' Wright Mediation Sensitivity Analysis
#' 
#' Computes the Minimum Detectable Effect Size (MDES) in terms of indirect R2
#' given a fixed sample size, returning the Best (Balanced) and Worst (Bottleneck) scenarios.
#' 
#' @param n Fixed sample size available for the study.
#' @param type "explore" (default), "consistent", or "suppression".
#' @param k_paths Number of multiplicative paths in the indirect block (default = 2).
#' @param c_prime Expected direct effect coefficient (default = 0.20).
#' @param r2_m_overlap Variance of the target mediator explained by other parallel mediators (VIF penalty).
#'        Tip: For 2 mediators with a correlation of 0.30, enter the squared value (0.30^2 = 0.09).
#'        Tip: For 3+ mediators, enter the multiple R2 of the target regressed on the others.
#' @param r2_parallel Variance of Y explained by other parallel mediators.
#' @param r2_covariates Variance of Y explained by external control variables.
#' @param power_target Desired joint statistical power threshold (default = 0.80).
#' @param alpha Significance level (default = 0.05).
#' @param max_beta Maximum allowed single standardized coefficient (guardrail, default = 0.80).

wright_sensitivity <- function(n, type = "explore", k_paths = 2, c_prime = 0.20,
                               r2_m_overlap = 0, r2_parallel = 0, r2_covariates = 0,
                               power_target = 0.80, alpha = 0.05, max_beta = 0.80) {
  
  # 1. INPUT VALIDATION & ARCHITECTURAL GUARDRAIL
  if(n < 10) stop("Sample size 'n' must be at least 10.")
  if(r2_covariates < 0 | r2_covariates >= 0.90) stop("r2_covariates must be between 0 and 0.90.")
  if(r2_parallel < 0 | r2_parallel >= 0.90) stop("r2_parallel must be between 0 and 0.90.")
  if(r2_m_overlap < 0 | r2_m_overlap >= 1) stop("r2_m_overlap must be strictly between 0 and 1.")
  if(k_paths < 2) stop("k_paths must be at least 2.")
  
  if(k_paths >= 4) {
    warning("Models with k >= 4 (3 or more serial mediators) are highly susceptible to severe variance attenuation and measurement error accumulation. Results are provided for mathematical evaluation, but empirical implementation requires extreme caution.", call. = FALSE)
  }
  
  # 2. PARALLEL MEDIATION PENALTY (Analytic VIF)
  vif_penalty <- 1 / (1 - r2_m_overlap)
  
  # 3. CORE SENSITIVITY ENGINE
  get_mdes <- function(cp) {
    r2_grid <- seq(0.0001, 0.8000, by = 0.0001)
    
    mdes_balanced   <- NA
    mdes_bottleneck <- NA
    
    for(r2 in r2_grid) {
      K <- sqrt(r2)
      min_beta <- K / (max_beta^(k_paths - 1))
      
      if(min_beta > max_beta) next
      
      # Compute residual variance of Y
      var_error_y <- 1 - (cp^2 + K^2 + 2 * cp * K) - r2_covariates - r2_parallel
      if(var_error_y <= 0) var_error_y <- 0.01
      
      # --- BALANCED SCENARIO ---
      beta_bal <- K^(1/k_paths)
      se_bal <- sqrt(var_error_y / (n * (1 - beta_bal^2))) * sqrt(vif_penalty)
      pow_bal <- pnorm((beta_bal / se_bal) - qnorm(1 - alpha/2))
      joint_pow_bal <- pow_bal^k_paths 
      
      # --- BOTTLENECK SCENARIO ---
      se_min <- sqrt(var_error_y / (n * (1 - min_beta^2))) * sqrt(vif_penalty)
      se_max <- sqrt(var_error_y / (n * (1 - max_beta^2))) * sqrt(vif_penalty)
      
      pow_min <- pnorm((min_beta / se_min) - qnorm(1 - alpha/2))
      pow_max <- pnorm((max_beta / se_max) - qnorm(1 - alpha/2))
      joint_pow_bot <- pow_min * (pow_max^(k_paths - 1))
      
      if(is.na(mdes_balanced) && joint_pow_bal >= power_target) {
        if(r2 == 0.0001) mdes_balanced <- "below_min" else mdes_balanced <- as.character(r2)
      }
      if(is.na(mdes_bottleneck) && joint_pow_bot >= power_target) {
        if(r2 == 0.0001) mdes_bottleneck <- "below_min" else mdes_bottleneck <- as.character(r2)
      }
      
      if(!is.na(mdes_balanced) && !is.na(mdes_bottleneck)) break
    }
    return(c(mdes_balanced, mdes_bottleneck))
  }
  
  # 4. OUTPUT GENERATION
  cat("\n==================================================================================\n")
  cat("                 WRIGHT MEDIATION SENSITIVITY ANALYSIS                 \n")
  cat("==================================================================================\n")
  cat(sprintf("Fixed Sample Size (N): %d \n", n))
  cat(sprintf("Structural Paths:      %d \n", k_paths))
  cat(sprintf("Direct Effect (c'):    %.2f (Base Magnitude)\n", abs(c_prime)))
  cat(sprintf("Other Mediators (R2):  %.2f (Variance in Y)\n", r2_parallel))
  cat(sprintf("Target M Overlap (R2): %.2f (Collinearity VIF = %.2f)\n", r2_m_overlap, vif_penalty))
  cat(sprintf("Covariates (R2):       %.2f \n", r2_covariates))
  cat(sprintf("Target Joint Power:    %.2f \n", power_target))
  cat("----------------------------------------------------------------------------------\n")
  cat(sprintf("%-12s %-12s %-18s %-18s %-12s\n", "SCENARIO", "MODEL TYPE", "IMPLIED MIN BETA", "IMPLIED MAX BETA", "REQUIRED R2"))
  cat("----------------------------------------------------------------------------------\n")
  
  scenarios <- if(type == "explore") c("consistent", "suppression") else type
  high_beta_warning <- FALSE
  
  for(t in scenarios) {
    cp_val <- if(t == "consistent") abs(c_prime) else -abs(c_prime)
    t_label <- if(t == "consistent") "Consistent" else "Suppression"
    
    results <- get_mdes(cp_val)
    mdes_bal <- results[1]
    mdes_bot <- results[2]
    
    
    if (is.na(mdes_bal)) {
      r2_bal_str <- "Unreachable"
      b_bal_min_str <- "-"
      b_bal_max_str <- "-"
    } else if (mdes_bal == "below_min") {
      r2_bal_str <- "< 0.0001"
      beta_bal_implied <- (0.01)^(1/k_paths)
      b_bal_min_str <- sprintf("%.3f", beta_bal_implied)
      b_bal_max_str <- sprintf("%.3f", beta_bal_implied)
      if(beta_bal_implied > 0.50) high_beta_warning <- TRUE
    } else {
      r2_bal_str <- sprintf("%.4f", as.numeric(mdes_bal))
      beta_bal_implied <- sqrt(as.numeric(mdes_bal))^(1/k_paths)
      b_bal_min_str <- sprintf("%.3f", beta_bal_implied)
      b_bal_max_str <- sprintf("%.3f", beta_bal_implied)
      if(beta_bal_implied > 0.50) high_beta_warning <- TRUE
    }
    
    
    if (is.na(mdes_bot)) {
      r2_bot_str <- "Unreachable"
      b_bot_min_str <- "-"
      b_bot_max_str <- "-"
    } else if (mdes_bot == "below_min") {
      r2_bot_str <- "< 0.0001"
      beta_bot_implied <- 0.01 / (max_beta^(k_paths - 1))
      b_bot_min_str <- sprintf("%.3f", beta_bot_implied)
      b_bot_max_str <- sprintf("%.3f (x%d)", max_beta, k_paths - 1)
      if(beta_bot_implied > 0.50) high_beta_warning <- TRUE
    } else {
      r2_bot_str <- sprintf("%.4f", as.numeric(mdes_bot))
      beta_bot_implied <- sqrt(as.numeric(mdes_bot)) / (max_beta^(k_paths - 1))
      b_bot_min_str <- sprintf("%.3f", beta_bot_implied)
      b_bot_max_str <- sprintf("%.3f (x%d)", max_beta, k_paths - 1)
      if(beta_bot_implied > 0.50) high_beta_warning <- TRUE
    }
    
    cat(sprintf("%-12s %-12s %-18s %-18s %-12s\n", "Balanced", t_label, b_bal_min_str, b_bal_max_str, r2_bal_str))
    cat(sprintf("%-12s %-12s %-18s %-18s %-12s\n", "Bottleneck", t_label, b_bot_min_str, b_bot_max_str, r2_bot_str))
    if(t != scenarios[length(scenarios)]) cat("- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -\n")
  }
  cat("==================================================================================\n")
  cat("Method: Joint Significance Power.\n\n")
  
  if(high_beta_warning) {
    warning("The estimated MDES requires a local step effect (implied beta) > 0.50. This is unusually large for social sciences.", call. = FALSE)
  }
}


################################################################################
###Function usage example!

#Single mediator model
wright_sensitivity(n = 107,
                   type = "explore",
                   k_paths = 2,
                   c_prime = 0.30,
                   r2_m_overlap = 0, 
                   r2_parallel = 0, 
                   r2_covariates = 0,
                   power_target = 0.80,
                   alpha = 0.05,
                   max_beta = 0.80)

#Two parallel mediators model
wright_sensitivity(n = 105, 
                   type = "explore", 
                   k_paths = 2,
                   c_prime = 0.30,
                   r2_m_overlap = 0.09,
                   r2_parallel = 0.09,
                   r2_covariates = 0,
                   power_target = 0.80, 
                   alpha = 0.05,
                   max_beta = 0.80)

#Two serial mediators model
wright_sensitivity(n = 136,
                   type = "explore",
                   k_paths = 3,
                   r2_m_overlap = 0, 
                   r2_parallel = 0, 
                   r2_covariates = 0,
                   power_target = 0.80,
                   alpha = 0.05,
                   max_beta = 0.80)

```

The implementation of the `wright_sensitivity` function thereby bridges the gap between available sample resources and statistical realism. However, applying these sensitivity and power metrics to multi-stage serial models introduces a fundamental mathematical paradox, requiring a deeper examination of how variance structurally behaves as the chain of paths multiply.

## 9. The Multiplicative Nature of Serial Mediation and Variance Erosion

In structural equation modeling and path analysis, the overall indirect effect of a serial mediation chain is calculated as the product of the individual path coefficients. When utilizing standardized coefficients ($\beta$), which typically range between -1.0 and 1.0, this multiplicative property inherently leads to a geometric attenuation of the serial mediation effect size. This mathematical necessity gives rise to a phenomenon that can be defined as "variance erosion."

The variance in the dependent variable explained specifically by the indirect effect — denoted as the indirect $R^2$ — is the square of this multiplied coefficient. For a serial mediation chain consisting of $k$ paths, assuming a mathematically balanced scenario where all local path coefficients are equal to $\beta$, the indirect variance is expressed as:

$$R^2_{\text{indirect}} = (\beta^k)^2 = \beta^{2k}$$

As the number of structural paths ($k$) increases, the resulting indirect $R^2$ decays exponentially. This decay occurs regardless of the substantive strength of the individual coefficients of each link.

To illustrate this mechanism, consider a three-path serial mediation model ($X \rightarrow M_1 \rightarrow M_2 \rightarrow Y$, where $k = 3$). Assume that each local step in the sequence demonstrates a standardized coefficient of $\beta = 0.30$, an effect size generally considered robust in the social and behavioral sciences. The standardized indirect effect is the product of these paths: $0.30 \times 0.30 \times 0.30 = 0.027$. Consequently, the actual variance explained by this entire indirect sequence is the square of the effect: $0.027^2 \approx 0.0007$.

From a macro-level perspective, an explained variance of 0.07% ($R^2 = 0.0007$) appears statistically trivial. This metric frequently causes a cognitive disconnect for researchers, leading to the erroneous conclusion that the theoretical model is weak or that the empirical test has failed. However, this global metric conceals the micro-level reality of the model: the structural relationships connecting the variables are substantively meaningful ($\beta = 0.30$), but their sequential multiplication inevitably produces a fractionally small overall parameter.

Therefore, when designing studies, conducting a priori power analyses, or interpreting sensitivity outputs for models with $k \ge 3$, researchers should evaluate the implied minimum beta of the local paths. Relying solely on the magnitude of the serial indirect $R^2$ may mislead the interpretation of the model's validity and statistical requirements.

To empirically demonstrate variance erosion and validate the analytical assumptions described above, we implement a structured synthetic data-generation process. The aim is to monitor the behavior of the serial indirect effect as the number of serial paths ($k$) progressively increases, while keeping both the sample size ($N$) and the magnitude of the individual structural coefficients ($\beta$) strictly constant.

The logic of this simulated demonstration is articulated in four sequential phases: 1. **Dataset Generation:** Creation of synthetic data matrices based on structural equation models with fixed standardized parameters ($\beta = 0.30$ for each local path, $c' = 0.20$ for the direct effect). 2. **Model Estimation:** Fitting the serial mediation models using the estimation algorithm provided by the `lavaan` package. 3. **Quantification of the Serial Indirect Effect Size (**$R^2_{indirect}$): Extraction of the standardized parameters and calculation of the variance explained by the serial effect. 4. **Critical Power/Sensitivity Analysis:** Evaluation of the discrepancy between the local significance of the individual segments ($'MIN and MAX betas'$) and the magnitude of the serial indirect effect $R^2_{\text{indirect}}$.

### 9.1 Simulated Processes and Empirical Evidence of Variance Erosion (point 1 to 3)

The following code chunk executes the comparative demonstration for three structural configurations: a simple mediation ($k=2$), a two-mediator serial chain ($k=3$), and a complex three-mediator chain ($k=4$).

```{r}
# ==============================================================================
# EMPIRICAL DEMONSTRATION: VARIANCE EROSION IN SERIAL MEDIATION
# ==============================================================================
library(lavaan)

# 1. DATA GENERATION FUNCTIONS (Manual Approach)
# ------------------------------------------------------------------------------
simulate_k2 <- function(N, b_val, cp, seed = 42) {
  set.seed(seed)
  X  <- rnorm(N, mean = 0, sd = 1)
  M1 <- b_val * X + rnorm(N, mean = 0, sd = 1)
  Y  <- cp * X + b_val * M1 + rnorm(N, mean = 0, sd = 1)
  return(data.frame(X, M1, Y))
}

simulate_k3 <- function(N, b_val, cp, seed = 42) {
  set.seed(seed)
  X  <- rnorm(N, mean = 0, sd = 1)
  M1 <- b_val * X + rnorm(N, mean = 0, sd = 1)
  M2 <- b_val * M1 + rnorm(N, mean = 0, sd = 1)
  Y  <- cp * X + b_val * M2 + rnorm(N, mean = 0, sd = 1)
  return(data.frame(X, M1, M2, Y))
}

simulate_k4 <- function(N, b_val, cp, seed = 42) {
  set.seed(seed)
  X  <- rnorm(N, mean = 0, sd = 1)
  M1 <- b_val * X + rnorm(N, mean = 0, sd = 1)
  M2 <- b_val * M1 + rnorm(N, mean = 0, sd = 1)
  M3 <- b_val * M2 + rnorm(N, mean = 0, sd = 1)
  Y  <- cp * X + b_val * M3 + rnorm(N, mean = 0, sd = 1)
  return(data.frame(X, M1, M2, M3, Y))
}

# 2. CREATION OF DATASETS (N = 500, Moderate local paths)
# ------------------------------------------------------------------------------
df_k2 <- simulate_k2(N = 500, b_val = 0.50, cp = 0.20)
df_k3 <- simulate_k3(N = 500, b_val = 0.50, cp = 0.20)
df_k4 <- simulate_k4(N = 500, b_val = 0.50, cp = 0.20)

# 3. LAVAAN MODEL DEFINITION (Labeling paths for r2_indirect)
# ------------------------------------------------------------------------------
mod_k2 <- "
  # Paths
  M1 ~ a * X
  Y  ~ cp * X + b * M1
"

mod_k3 <- "
  # Paths
  M1 ~ a * X
  M2 ~ b * M1
  Y  ~ cp * X + c * M2
"

mod_k4 <- "
  # Paths
  M1 ~ a * X
  M2 ~ b * M1
  M3 ~ c * M2
  Y  ~ cp * X + d * M3
"

fit_k2 <- sem(mod_k2, estimator = "MLR", data = df_k2)
fit_k3 <- sem(mod_k3, estimator = "MLR", data = df_k3)
fit_k4 <- sem(mod_k4, estimator = "MLR", data = df_k4)

summary(fit_k2, standardized = TRUE)
summary(fit_k3, standardized = TRUE)
summary(fit_k4, standardized = TRUE)

# ==============================================================================
# RUNNING r2_indirect ON RAW COEFFICIENTS
# ==============================================================================

cat("\n=== LAUNCHING r2_indirect ON RAW UNSTANDARDIZED DATA ===\n")
# Using a small number of bootstrap resamples (n_boot = 200) for fast compilation in the vignette
invisible(capture.output({
res_k2 <- r2_indirect(fit_k2, x_var = "X", y_var = "Y", Med_Path = c("a", "b"), n_boot = 200)
res_k3 <- r2_indirect(fit_k3, x_var = "X", y_var = "Y", Med_Path = c("a", "b", "c"), n_boot = 200)
res_k4 <- r2_indirect(fit_k4, x_var = "X", y_var = "Y", Med_Path = c("a", "b", "c", "d"), n_boot = 200)
}))

cat("\n--> FUNCTION REPORT: K = 2 (Simple Mediation)\n")
print(res_k2, row.names = FALSE)

cat("\n--> FUNCTION REPORT: K = 3 (Two Serial Mediators)\n")
print(res_k3, row.names = FALSE)

cat("\n--> FUNCTION REPORT: K = 4 (Three Serial Mediators)\n")
print(res_k4, row.names = FALSE)

```

The outputs generated by these simulated datasets provide concrete empirical evidence of variance erosion. Despite holding the local structural parameters entirely constant ($\beta = 0.50$), the mere addition of sequential mediation steps geometrically diminishes the measurable serial indirect effect size. This mathematical inevitability sets the stage for a critical methodological trap in sample size planning.

### 9.2 Showing the Power/Sensitivity Analysis Paradox in Serial Models

The empirical demonstration above highlights a critical methodological trap: as the number of paths ($k$) increases, the serial indirect $R^2$ drastically erodes, even when the underlying local structural paths remain robust.

TThis mathematical artifact creates a conceptual "power analysis paradox." Because the global indirect variance of a complex serial chain (e.g., $R^2 = 0.001$ for $k = 4$) decays exponentially, evaluating the model strictly through this global metric creates the optical illusion of an empirically trivial effect.

However, it should be acknowledged that the interpretation of the magnitude of the serial indirect effect in terms of explained variance warrants further methodological investigation to establish appropriate effect-size conventions (and, more generally, for indirect effects). While an indirect $R^2$ of 0.0001 might currently be dismissed as negligible, such judgments rely on traditional benchmarks that are ill-suited for complex structural models. Currently, there is a lack of rigorous meta-analytical research tailored to the inherently attenuated nature of serial chains, which is necessary to properly define what constitutes a "small," "medium," or "large" serial indirect effect.

Evaluating statistical power for complex serial indirect effects should be based on the following logic underlying Joint Significance Testing. In structural equation modeling, a serial indirect effect is deemed significant if, and only if, **all** of its constituent local paths are statistically significant. Therefore, the actual statistical power of a complex serial model is not dictated by the potential infinitesimal indirect $R^2$, but rather by the sample size required to reliably detect the *weakest* individual path in the chain (the bottleneck).

To properly resolve this paradox, the `wright_power()` and `wright_sensitivity()` functions introduce a targeted framework for complex serial mediation (i.e., $k > 3$). Rather than naively relying on the eroded indirect $R^2$, these algorithms allow researchers to anchor their analytical planning directly to the **implied minimum beta** of the local paths. This ensures that sample size and sensitivity estimations are perfectly aligned with the actual structural mechanics of the model.

### 9.2.1 The Price of Complexity in a-priori power analysis for serial mediation models: Local vs. "Global" Perspectives

When scaling up a serial mediation chain from *k* = 2 to *k* = 4 structural paths, researchers face a mathematical trade-off between **local mechanism strength** and **"serial" explanatory power**. To fully understand how variance attenuation affects a study design, the `WrightMediation` engine allows you to simulate power from two distinct perspectives:

------------------------------------------------------------------------

### Scenario A: The Global Perspective (Fixed Indirect Effect of the serial chain)

In this scenario, we mandate that the serial indirect chain must explain a meaningful, fixed proportion of variance in the final outcome *Y* (e.g., *R*² = 0.015, or 1.5%), regardless of how many mediators are packed in between.

- **The Diagnostic Goal:** Observe what magnitude the intermediate step-effects must reach to mathematically sustain that "serial" variance.
- **The Insight:** As *k* increases, the function forces the local path coefficients to increase. By *k* = 4, the required intermediate betas will overshoot realistic social-science thresholds (exceeding 0.50), automatically triggering the package's methodological guardrails and warnings.

```{r}
# -----------------------------------------------------------------------
# SCENARIO A: The Global Perspective (Fixed Indirect R2 = 0.015)
# -----------------------------------------------------------------------
cat("\n=======================================================================\n")
cat("     SCENARIO A: FIXED GLOBAL INDIRECT VARIANCE (r2_indirect = 0.015)   \n")
cat("=======================================================================\n")
# We set a theoretical expectation for the serial indirect variance (r2_indirect = 0.015)
# and compare the required sample size and implied step-effects across k = 2, 3, and 4.
# Notice how the implied local betas increase to mathematically sustain the serial R2,
# forcing N to adjust and automatically triggering the package's architectural guardrails.

cat("\n--> POWER ANALYSIS: K = 2 (Simple Mediation)\n")
pow_k2 <- wright_power(r2_indirect = 0.015, k_paths = 2, alpha = 0.05, power = 0.80)

cat("\n--> POWER ANALYSIS: K = 3 (Two Serial Mediators)\n")
pow_k3 <- wright_power(r2_indirect = 0.015, k_paths = 3, alpha = 0.05, power = 0.80)

cat("\n--> POWER ANALYSIS: K = 4 (Three Serial Mediators)\n")
pow_k4 <- wright_power(r2_indirect = 0.015, k_paths = 4, alpha = 0.05, power = 0.80)
```

------------------------------------------------------------------------

### Scenario B: The Local Perspective (Fixed Step-by-Step Effects)

In this scenario, we assume that each intermediate mechanism has a stable, moderate effect size (e.g., a standardized path coefficient of β = 0.35 for every step in the chain).

- **The Diagnostic Goal:** Observe what happens to the serial variance explained in *Y* as the chain grows.
- **The Insight:** You will notice that while the required sample size (*N*) remains relatively stable (because local power is driven by the single path magnitude), the Indirect *R*² drops exponentially. By *k* = 4, the total indirect effect vanishes into statistical background noise, rendering the overall model empirically trivial despite having "powered" local paths.

```{r}
# -----------------------------------------------------------------------
# SCENARIO B: The Local Perspective (Fixed Path Beta = 0.35)
# -----------------------------------------------------------------------
cat("\n=======================================================================\n")
cat("   SCENARIO B: FIXED LOCAL PATH COEFFICIENTS (target_beta_path = 0.35)  \n")
cat("=======================================================================\n")
# We set a theoretical expectation for the local path (target_beta_path = 0.35)
# and compare the required sample size across k = 2, 3, and 4.
# Notice how the required N remains stable because it is driven by the local effect,
# even though the translated serial R2 drops significantly due to variance attenuation.

cat("\n--> POWER ANALYSIS: K = 2 (Simple Mediation)\n")
pow_k2 <- wright_power(target_beta_path = 0.35, k_paths = 2, alpha = 0.05, power = 0.80)

cat("\n--> POWER ANALYSIS: K = 3 (Two Serial Mediators)\n")
pow_k3 <- wright_power(target_beta_path = 0.35, k_paths = 3, alpha = 0.05, power = 0.80)

cat("\n--> POWER ANALYSIS: K = 4 (Three Serial Mediators)\n")
pow_k4 <- wright_power(target_beta_path = 0.35, k_paths = 4, alpha = 0.05, power = 0.80)
```

These comparative scenarios expose the power analysis paradox. They mathematically demonstrate that researchers must anchor their statistical planning to the implied local mechanisms (Scenario B) rather than a serial variance target (Scenario A).

### 9.2.2 Sensitivity Profiles: Standard vs. Boundary Perspectives

When running a sensitivity analysis using `wright_sensitivity` with `type = "explore"`, the sample size (*N*) is held constant, and the function calculates the Minimum Detectable Effect Size (MDES) — expressed as both the Required Serial Indirect *R*² and the Implied Local Betas — across both **Consistent** and **Suppression** mediation configurations simultaneously. To understand how model complexity interacts with sample size, we evaluate sensitivity from three operational perspectives:

------------------------------------------------------------------------

### Scenario A: The Standard Sample Perspective (Fixed N = 800)

In this scenario, we evaluate a realistic, well-powered sample size across different chain lengths (*k* = 2, 3, and 4) to map out both structural behaviors side by side.

- **The Diagnostic Goal:** Observe how the minimum detectable thresholds shift as the mediation chain grows, contrasting how the directionality of the direct effect (coherence vs. suppression) impacts the required parameters.
- **The Insight:** You will notice that as *k* increases, even if the required serial *R*² remains in a similar ballpark (though incrementally decreasing), the implied intermediate betas must scale up significantly to ensure the joint significance of all paths.

```{r}
# -----------------------------------------------------------------------
# SCENARIO A: The Standard Sample Perspective (Fixed N = 800)
# -----------------------------------------------------------------------
cat("\n=======================================================================\n")
cat("        SCENARIO A: STANDARD SAMPLE SENSITIVITY (Fixed n = 800)         \n")
cat("=======================================================================\n")

# We set a fixed sample size for a standard study (n = 800)
# and evaluate the Minimum Detectable Effect Size (MDES) across k = 2, 3, and 4.
# Notice how the required serial R2 and implied local betas adjust to maintain target power,
# illustrating the baseline sensitivity of standard empirical designs.

cat("\n>>> Sensitivity for k = 2 <<<\n")
wright_sensitivity(n = 800, k_paths = 2, type = "explore")

cat("\n>>> Sensitivity for k = 3 <<<\n")
wright_sensitivity(n = 800, k_paths = 3, type = "explore")

cat("\n>>> Sensitivity for k = 4 <<<\n")
wright_sensitivity(n = 800, k_paths = 4, type = "explore")
```

------------------------------------------------------------------------

### Scenario B: The Large Sample Perspective (Fixed N = 50,000)

In this scenario, we push the sample size to an extreme high-power condition to stress-test the empirical limits of the model lengths.

- **The Diagnostic Goal:** Test the mathematical resolution and the lower boundaries of the sensitivity engine under both exploratory configurations.
- **The Insight:** Because the sample is massive, the true minimum detectable *R*² drops below the structural grid floor. This safely triggers the package's `< 0.0001` protective guardrail in the output table for both Consistent and Suppression rows, while seamlessly maintaining the exact mathematical calculation for the implied local path coefficients.

```{r}
# -----------------------------------------------------------------------
# SCENARIO B: The Large Sample Perspective (Fixed N = 50,000)
# -----------------------------------------------------------------------
cat("\n=======================================================================\n")
cat("      SCENARIO B: LARGE SAMPLE BOUNDARY SENSITIVITY (Fixed n = 50000)   \n")
cat("=======================================================================\n")

# We set a massive sample size (n = 50000) to test the limits of the sensitivity grid
# and evaluate the required serial R2 and implied local betas across k = 2, 3, and 4.
# Notice how the required R2 drops below the 0.0001 threshold, successfully triggering
# the \"< 0.0001\" protective guardrail without breaking the implied beta computation.

cat("\n>>> Sensitivity for k = 2 <<<\n")
wright_sensitivity(n = 50000, k_paths = 2, type = "explore")

cat("\n>>> Sensitivity for k = 3 <<<\n")
wright_sensitivity(n = 50000, k_paths = 3, type = "explore")

cat("\n>>> Sensitivity for k = 4 <<<\n")
wright_sensitivity(n = 50000, k_paths = 4, type = "explore")
```

### Scenario C: The Underpowered Small Sample Perspective (Fixed N = 150)

In this scenario, we evaluate a highly restricted, underpowered sample size across the same chain lengths ($k$ = 2, 3, and 4) to observe the mathematical limits of empirical feasibility.

- **The Diagnostic Goal:** Map out the inflation of the Minimum Detectable Effect Size (MDES) when the study design is severely sample-constrained, tracking the strict requirements forced upon individual structural segments.
- **The Insight:** You will notice that because the sample size is small, the function is forced to demand an exceptionally high implied minimum beta for the local paths to satisfy the joint significance criteria. As structural complexity scales up, these required local coefficients increase to magnitudes that are often boundaries-defying or substantively implausible in the behavioral sciences, explicitly illustrating why complex serial mechanisms cannot be reliably detected with underpowered designs.

```{r}
# -----------------------------------------------------------------------
# SCENARIO C: The Underpowered Small Sample Perspective (Fixed N = 150)
# -----------------------------------------------------------------------
cat("\n=======================================================================\n")
cat("      SCENARIO C: SMALL SAMPLE UNDERPOWERED SENSITIVITY (Fixed n = 150) \n")
cat("=======================================================================\n")

# We set a restricted sample size for an underpowered study (n = 150)
# and evaluate he required serial R2 and implied local betas across k = 2, 3, and 4.
# Notice how the implied local betas increase to exceptionally high values,
# illustrating the severe empirical constraints of small-sample serial designs.

cat("\n>>> Sensitivity for k = 2 <<<\n")
wright_sensitivity(n = 150, k_paths = 2, type = "explore")

cat("\n>>> Sensitivity for k = 3 <<<\n")
wright_sensitivity(n = 150, k_paths = 3, type = "explore")

cat("\n>>> Sensitivity for k = 4 <<<\n")
wright_sensitivity(n = 150, k_paths = 4, type = "explore")
```

This comprehensive evaluation across low, standard, and extreme sample size boundaries validates the structural integrity of the framework. By illustrating the mathematical realities of variance erosion and providing precise diagnostic tools for both prospective power and retrospective sensitivity, this analytical approach ensures that researchers may confidently design, evaluate, and interpret simple mediation (Single, Parallel and Two serial mediation models) on the base of $R^2_{\text{indirect}}$, while in the context of complex serial mediation models on the base of local effect size (i.e., $MIN$ and $MAX$ implied betas), while remaining strictly within the bounds of empirical plausibility.
